---
title: "Data Preperation"
output: html_document
date: "2025-06-30"
---

# Summary

## Transactions

Reads the NSW Valuer-General sales file dated 2025-06-30.

-   **Address normalisation** – concatenates unit/house number, street name, locality and postcode into a single lowercase string; trims extra spaces and standardises street-type suffixes with `type_map`.
-   Produces a clean `transactions` table ready for matching to the national address file.

## GNAF

Imports the August 2024 NSW edition of G-NAF (`ADDRESS_DETAIL`, `STREET_LOCALITY`, `LOCALITY`).

-   Reconstructs a textual address in exactly the same format used for transactions, then attaches the unique `ADDRESS_DETAIL_PID`.
-   Cleans long street types to their short forms so the two datasets align perfectly.

## Merge GNAF and Transactions

Tests the join—≈ 95 % of sales match a G-NAF PID. Keeps only matched rows and restricts to contracts from 2022-01-01 onward with `Primary purpose == "Residence"`. Saves an intermediate CSV for backup.

## GEO Codes

Pulls longitude and latitude for every PID from `NSW_ADDRESS_DEFAULT_GEOCODE` and converts the transactions into an `sf` POINT layer (`transactions_sf`) in EPSG 7844. A random 10 000-point plot acts as a visual QA step.

## Statistical Areas, Postcodes and Amenities

-   Reads SA1, SA2 and POA polygons and pre-computes their centroids.
-   Loads point layers for hospitals, parks and rail stops.
-   Defines the Sydney CBD as the union of two SA2 polygons and extracts its centroid for distance work.

## Shops (from G-NAF)

Filters the G-NAF detail table for records where `FLAT_TYPE_CODE == "SHOP"`, joins their default geocodes and converts the result to an `sf` POINT layer called `shops`.

## Calculate Geographic Variables

Defines `add_geo_vars()`, a reusable helper that:

1.  Re-projects, validates and converts polygons to centroids.
2.  Computes straight-line distances from each centroid (or point) to the CBD, nearest hospital, park and rail stop.
3.  Counts G-NAF-coded shops within 1 km, 2 km and 5 km buffers.

The helper is applied once to SA1, SA2 and POA layers, then to the much larger `transactions_sf` in ten equal-sized chunks to avoid memory spikes; the chunk results are row-bound into a single enriched transactions layer.

## Merge All

Uses `add_area_vars_optimized()` to attach the new accessibility and amenity variables from SA1 (`…_sa1`), SA2 (`…_sa2`) and POA (`…_poa`) back onto each transaction point. The join is a two-step process—lean spatial lookup for the area code followed by a fast attribute merge—so only the necessary columns are added.

## Save File

Writes the prepared transaction GeoPackage under `Data/Intermediate`, then
adds the 2026 regional attributes and writes the final analysis GeoPackage
under `Data/Analysis`.

## Additional Attributes (2026)

Updates in year 2026. Create additional attributes for each geographical unit (SA1, SA2 and POA) to include:

1.  `sydney` = indicator variable if within greater Sydney boundary (ABS ASGS Ed. 3)

2.  `shr_resid,shr_dwell` = share (%) of area that is zoned residential (R) or any zoning layer that permits dwellings (per NSW Zoning & Planning Layers)

3.  `var_circ` = measure of circularity, where 0 = perfectly circular. Obtained by:

    1.  Fitting a circle using least-squares to the centroid of geog. unit: $\min_{r} \sum_{i=1}^{n} \left( \sqrt{(x_i - x_0)^2 + (y_i - y_0)^2} - r \right)^2(x_0​,y_0​,r)$ where $(x_0, y_0)$ is the center of the circle, $(x_i,y_i)$ is the $i$th boundary point of a geog. unit, and $r$ is the radius.

    2.  Calculating the area that falls outside of the circle plus where the circle does not overlap w/ geog. unit (numerator)

    3.  Evaluating as a share of the total area of the geog. unit (denominator)

4.  Calculate the Hopkins-Skellam statistic $A_k$ for each set of transactions per geog. unit $k$, setting the geog. unit as the window boundary. Monte-Carlo with 999 simulations. Statistic is calculated for each below:

    $A_k=\frac{\sum^m_{i=1}d_i^2}{\sum^m_{j=1}e_j^2}$

    where $m$ is the number of sampled points (999 in this case), $d_i$ are nearest-neighbour distances, and $e_j=d(u_j)$ empty-space distances for equal number of $m$ randomly sampled spatial locations.

# Packages

Run the notebook from the `Code` directory. All input and output paths below
point to the project-level `Data` directory through `../Data`.

```{r}
library(dplyr)
library(data.table)
library(stringr)
library(ggplot2)
library(sf)
library(purrr)
library(progressr)
```

# Transactions

```{r}

transactions = fread('../Data/Inputs/nsw-property-sales-data-updated20250630.csv')
# ../Data/Inputs/GNAF/NSW_ADDRESS_DETAIL_psv.psv
# Merge string of `Property unit number`, `Property house number`, `Property street name`, `Property locality`, `Property postcode` into a single address string

transactions = transactions %>%
  mutate(
    address = ifelse(
      (is.na(`Property unit number`) | (`Property unit number` == '')),
      paste(
        str_replace(`Property house number`, ' ', ''),
        `Property street name`,
        `Property locality`,
        `Property post code`,
        sep = ','
      ),
      paste(
        str_replace(`Property unit number`, ' ', ''),
        str_replace(`Property house number`, ' ', ''),
        `Property street name`,
        `Property locality`,
        `Property post code`,
        sep = ','
      )
    ),
    address = tolower(address)  # Convert address to lowercase
  ) %>%
  filter(!str_starts(address, ','))
type_map <- c(
  "street,"    = "st,",
  "road,"      = "rd,",
  "avenue,"    = "ave,",
  "drive,"     = "dr,",
  "boulevard," = "bvd,",
  "place,"     = "pl,",
  "close,"     = "cl,",
  "crescent,"  = "cres,",
  "court,"     = "ct,",
  "circuit,"   = "cct,",
  "parade,"    = "pde,",
  "highway,"   = "hwy,",
  "lane,"      = "ln,",
  "terrace,"   = "tce,",
  "parkway,"   = "pkwy,"
)

# Replace street types in the address
for (type in names(type_map)) {
  transactions$address <- str_replace_all(transactions$address, fixed(type), type_map[type])
}

head(transactions$address, 50)
```

# GNAF

```{r}

# 1.  load the three NSW tables ---------------------------------------------
detail  <- fread("../Data/Inputs/GNAF/NSW_ADDRESS_DETAIL_psv.psv",  sep = "|")
street  <- fread("../Data/Inputs/GNAF/NSW_STREET_LOCALITY_psv.psv", sep = "|")
local   <- fread("../Data/Inputs/GNAF/NSW_LOCALITY_psv.psv",        sep = "|")

# ── look-ups ──────────────────────────────────────────────────
street_key <- street %>% 
  transmute(
    STREET_LOCALITY_PID,
    street_name = str_squish(str_to_lower(STREET_NAME)),
    street_type = str_squish(str_to_lower(STREET_TYPE_CODE))   # rd, st, ave …
  )

local_key <- local %>% 
  transmute(
    LOCALITY_PID,
    locality = str_squish(str_to_lower(LOCALITY_NAME))
  )

# ── build address exactly like transactions ──────────────────
gnaf_with_addr <- detail %>% 
  left_join(street_key, by = "STREET_LOCALITY_PID") %>% 
  left_join(local_key,  by = "LOCALITY_PID") %>% 
  mutate(
    flat_num  = str_remove_all(FLAT_NUMBER,      "\\s+"),
    house_num = paste0(
                  str_remove_all(NUMBER_FIRST,   "\\s+"),
                  tolower(coalesce(str_remove_all(NUMBER_FIRST_SUFFIX, "\\s+"), ""))
                ),
    street    = str_trim(paste(street_name, street_type)),

    address = case_when(
      !is.na(flat_num) & flat_num != "" ~
        paste(flat_num, house_num, street, locality, POSTCODE, sep = ","),
      TRUE ~ paste(house_num, street, locality, POSTCODE, sep = ",")
    ) |> str_squish() |> str_to_lower()
  ) %>% 
  filter(!str_starts(address, ",")) %>% 
  select(ADDRESS_DETAIL_PID, address)

# Replace long with short names

gnaf_with_addr$address <- str_replace_all(
  gnaf_with_addr$address,
  type_map          # just hand the named vector straight in
)

# sanity-check
head(gnaf_with_addr$address, 20)


rm()
```

# Merge GNAF and Transactions

Let's see the percentage of transactions we can map into the GNAF:

```{r}
mean(transactions$address %in% gnaf_with_addr$address, na.rm = TRUE) * 100
```

Almost 95%. This is sufficiently good for the time being. So, I will drop the rest of the observations from our sample. It can be further improved later

```{r}
transactions = merge(transactions, gnaf_with_addr, by = 'address', all.x = F, all.y = F)
#let's clean up the moemory!
rm(gnaf_with_addr, street, local, detail, street_key, local_key)
gc()
```

I will also drop observations before 2022. This leaves us with \# observations:

```{r}
transactions = transactions %>%
  filter(as.Date(`Contract date`) >= '2022-01-01') %>%
  filter(`Primary purpose` == 'Residence')
nrow(transactions)
```

Let's save this just in case

```{r}
fwrite(transactions, '../Data/Intermediate/transactions_with_GNAF-ID_2022_2025.csv')
# fread('../Data/Intermediate/transactions_with_GNAF-ID_2022_2025.csv')
```

# GEO Codes

```{R}
# 1. Load the NSW default-geocode table (this is where G-NAF stores the coords)
geo <- fread(
  "../Data/Inputs/GNAF/NSW_ADDRESS_DEFAULT_GEOCODE_psv.psv",
  sep    = "|",
  select = c("ADDRESS_DETAIL_PID", "LONGITUDE", "LATITUDE")
)

# 2. Attach longitude & latitude to the transactions
transactions <- transactions %>% 
  left_join(geo, by = "ADDRESS_DETAIL_PID") 

# 3. (Optional) convert to an sf object for spatial work
transactions_sf <- st_as_sf(
  transactions,
  coords = c("LONGITUDE", "LATITUDE"),
  crs    = 7844
)

rm(geo)
gc()

```

Visualize some random 10000 trsnactions

```{r}
ggplot() +
  geom_sf(data = transactions_sf[sample(nrow(transactions_sf), 10000), ], 
          aes(), size = 0.5, alpha = 0.5) +
  labs(title = "Random Sample of NSW Property Transactions",
       x = "Longitude", y = "Latitude") +
  theme_minimal() 
```

# Statistical areas, postcodes and amenities

Let's get amenities

```{r}
all_sa1<- st_read('../Data/Inputs/Boundaries/sa1_all.gpkg') %>% filter(str_starts(SA1_CODE21, '1'))
st_crs(all_sa1) <- 7844
all_sa1 <- all_sa1 %>% mutate(centroid = st_centroid(geom)) #st_union()
  
all_sa2 <- st_read('../Data/Inputs/Boundaries/sa2_all.gpkg') %>% filter(str_starts(SA2_CODE21, '1'))
st_crs(all_sa2) <- 7844
all_sa2 <- all_sa2 %>% mutate(centroid = st_centroid(geom))

all_poa <- st_read('../Data/Inputs/Boundaries/poa_all.gpkg') %>% filter(str_starts(POA_CODE21, '2'))
st_crs(all_poa) <- 7844
all_poa <- all_poa %>% mutate(centroid = st_centroid(geom))
```

```{r}
nsw_hosp <- st_read('../Data/Inputs/Amenities/nsw_hospital.gpkg')
st_crs(nsw_hosp) <- 7844
nsw_park <- st_read('../Data/Inputs/Amenities/nsw_park.gpkg')
st_crs(nsw_park) <- 7844
nsw_rail<- st_read('../Data/Inputs/Amenities/nsw_railstop.gpkg')
st_crs(nsw_rail) <- 7844

nsw_cbd <- all_sa2 %>% 
  filter(str_detect(SA2_NAME21,
                    regex("Sydney",
                          ignore_case = TRUE))) %>%
  filter(SA2_CODE21 == 117031644 | SA2_CODE21 == 117031645)
# merge polygons: 
nsw_cbd <- st_union(nsw_cbd)
# find cetroid
nsw_cbd_centroid <- st_centroid(nsw_cbd)
```

Get the shops and their coordinates from gNAF

```{r}
# 1.  Reload the NSW detail table with minimal columns
shop_cols <- c(
  "ADDRESS_DETAIL_PID",
  "FLAT_TYPE_CODE",
  "FLAT_NUMBER",           # optional: can help with QA
  "NUMBER_FIRST", "NUMBER_FIRST_SUFFIX",
  "STREET_LOCALITY_PID",   # optional: join to street table later if desired
  "LOCALITY_PID",          # "
  "POSTCODE"
)

detail_shop <- fread(
  "../Data/Inputs/GNAF/NSW_ADDRESS_DETAIL_psv.psv",
  sep    = "|",
  select = shop_cols
) %>% 
  filter(toupper(FLAT_TYPE_CODE) == "SHOP")    # keep only shops

# 2.  Pull the default geocode for every address
geo_shop <- fread(
  "../Data/Inputs/GNAF/NSW_ADDRESS_DEFAULT_GEOCODE_psv.psv",
  sep    = "|",
  select = c("ADDRESS_DETAIL_PID", "LONGITUDE", "LATITUDE")
)

# 3.  Merge → one row per shop with coordinates
shops <- detail_shop %>% 
  left_join(geo_shop, by = "ADDRESS_DETAIL_PID") 

rm(geo_shop)
# 4.  How many shops in NSW (G-NAF August 2024)?
nrow(shops)
```

Convert to shapefile

```{r}
shops  <- st_as_sf(shops, coords = c("LONGITUDE", "LATITUDE"), crs = 7844)
```

# GNAF Centroids

```{r}
# --------------------------------------------------------------------------
# GNAF Centroids  ·  residential stock only, split by dwelling type
# --------------------------------------------------------------------------

handlers("txtprogressbar")

# ── 1 · Load coordinates ----------------------------------------------------
gnaf_geo <- fread(
  "../Data/Inputs/GNAF/NSW_ADDRESS_DEFAULT_GEOCODE_psv.psv",
  sep    = "|",
  select = c("ADDRESS_DETAIL_PID", "LONGITUDE", "LATITUDE")
)

# ── 2 · Bring in flags & filter to residential stock -----------------------
detail_flags <- fread(
  "../Data/Inputs/GNAF/NSW_ADDRESS_DETAIL_psv.psv",
  sep    = "|",
  select = c("ADDRESS_DETAIL_PID", "FLAT_TYPE_CODE", "ALIAS_PRINCIPAL")
)

# commercial / service FLAT_TYPE_CODEs to exclude completely
commercial_codes <- c(
  "SHOP","OFFC","FCTY","WHSE","SHED","GRGE","CARS","KSK",
  "STOR","SEC","SITE","MBTH","WARD","STLL"
)

# dwelling-type buckets
house_codes <- c("", "HSE","TNHS","VLLA","DUPL","CTGE","BTSD","REAR")
apt_codes   <- c("UNIT","APT","FLAT","BLCK","PTHS","ROOM","STU","LOFT","MSNT","SE")

gnaf_geo <- gnaf_geo %>% 
  left_join(detail_flags, by = "ADDRESS_DETAIL_PID") %>% 
  filter(
    ALIAS_PRINCIPAL == "P",                                     # primary addr
    !(FLAT_TYPE_CODE %in% commercial_codes)                     # drop commercial
  ) %>% 
  mutate(
    FLAT_TYPE_CODE = tidyr::replace_na(FLAT_TYPE_CODE, ""),     # blank → ""
    dwelling_type  = case_when(
      FLAT_TYPE_CODE %in% house_codes ~ "house",
      FLAT_TYPE_CODE %in% apt_codes   ~ "apartment",
      TRUE                            ~ NA_character_
    )
  ) %>% 
  filter(!is.na(dwelling_type)) %>%                             # keep two buckets
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 7844)

# Show how many address records share coordinates. Address records, rather
# than unique coordinate locations, are the weights used below.
gnaf_coords <- st_coordinates(gnaf_geo)
centroid_check <- gnaf_geo %>%
  st_drop_geometry() %>%
  mutate(X = gnaf_coords[, "X"], Y = gnaf_coords[, "Y"]) %>%
  group_by(dwelling_type) %>%
  summarise(
    address_records = n(),
    unique_coordinates = n_distinct(X, Y),
    .groups = "drop"
  )
print(centroid_check)

rm(detail_flags, gnaf_coords, centroid_check); gc()

# ── 3 · Attach SA1 / SA2 / POA codes ---------------------------------------
gnaf_geo <- gnaf_geo %>% 
  st_join(all_sa1 %>% select(SA1_CODE21, SA2_CODE21), join = st_within) %>% 
  st_join(all_poa %>% select(POA_CODE21), join = st_within)

# ── 4 · Helper: centroid per area & dwelling bucket ------------------------
make_centroids <- function(sf_pts, area_col, dw_type) {
  sf_pts %>% 
    filter(dwelling_type == dw_type, !is.na(.data[[area_col]])) %>% 
    group_by(across(all_of(area_col))) %>% 
    summarise(
      geometry = st_centroid(st_combine(geometry)),
      .groups  = "drop"
    )
}

# ── 5 · Build & save the six layers ----------------------------------------
progressr::with_progress({
  p <- progressr::progressor(6)
  
  p("SA1 house");      gnaf_sa1_house <- make_centroids(gnaf_geo, "SA1_CODE21", "house")
  p("SA1 apartment");  gnaf_sa1_apt   <- make_centroids(gnaf_geo, "SA1_CODE21", "apartment")
  
  p("SA2 house");      gnaf_sa2_house <- make_centroids(gnaf_geo, "SA2_CODE21", "house")
  p("SA2 apartment");  gnaf_sa2_apt   <- make_centroids(gnaf_geo, "SA2_CODE21", "apartment")
  
  p("POA house");      gnaf_poa_house <- make_centroids(gnaf_geo, "POA_CODE21", "house")
  p("POA apartment");  gnaf_poa_apt   <- make_centroids(gnaf_geo, "POA_CODE21", "apartment")
})

st_write(gnaf_sa1_house, "../Data/Intermediate/gnaf_sa1_house_centroid.gpkg",
         delete_dsn = TRUE, quiet = TRUE)
st_write(gnaf_sa1_apt,   "../Data/Intermediate/gnaf_sa1_apartment_centroid.gpkg",
         delete_dsn = TRUE, quiet = TRUE)

st_write(gnaf_sa2_house, "../Data/Intermediate/gnaf_sa2_house_centroid.gpkg",
         delete_dsn = TRUE, quiet = TRUE)
st_write(gnaf_sa2_apt,   "../Data/Intermediate/gnaf_sa2_apartment_centroid.gpkg",
         delete_dsn = TRUE, quiet = TRUE)

st_write(gnaf_poa_house, "../Data/Intermediate/gnaf_poa_house_centroid.gpkg",
         delete_dsn = TRUE, quiet = TRUE)
st_write(gnaf_poa_apt,   "../Data/Intermediate/gnaf_poa_apartment_centroid.gpkg",
         delete_dsn = TRUE, quiet = TRUE)

```

# Calculate geographic variables

These must be calculated for each SA1, SA2, POA centroid.

-   distance to cbd
-   distance to nearest hospital
-   distance to nearest park
-   distance to nearest rail stop
-   the number shops within 1, 2 and 5 km distances.

```{R}
add_geo_vars <- function(
    area_sf,
    shops_sf       = shops,
    cbd_centroid   = nsw_cbd_centroid,
    hosp_sf        = nsw_hosp,
    park_sf        = nsw_park,
    rail_sf        = nsw_rail
) {

  # -------- Utilities (unchanged) -------------------------------------------
  crs_target <- 7844
  to_target  <- function(x) st_transform(x, crs = crs_target)

  prep_layer <- function(x) {
    x <- to_target(x)
    if (!all(st_is_valid(x))) x <- st_make_valid(x)
    st_centroid(x)
  }

  nearest_dist <- function(from_pts, to_pts) {
    nearest_idx <- st_nearest_feature(from_pts, to_pts)
    st_distance(from_pts, to_pts[nearest_idx, ], by_element = TRUE)
  }

  shop_count <- function(from_pts, shops_pts, r) {
    within_dist_list <- st_is_within_distance(from_pts, shops_pts, dist = r)
    lengths(within_dist_list)
  }

  # -------- Prepare Inputs (unchanged) --------------------------------------
  if ("centroid" %in% names(area_sf) && inherits(area_sf$centroid, "sfc")) {
    message("Using pre-existing 'centroid' column for calculations.")
    area_centroids <- to_target(area_sf$centroid)
  } else {
    message("No valid 'centroid' column found. Calculating centroids from main geometry.")
    area_centroids <- st_centroid(to_target(area_sf))
  }

  shops_sf     <- prep_layer(shops_sf)
  hosp_sf      <- prep_layer(hosp_sf)
  park_sf      <- prep_layer(park_sf)
  rail_sf      <- prep_layer(rail_sf)
  cbd_centroid <- prep_layer(cbd_centroid)

  # -------- Heavy work with progress bar (unchanged) ------------------------
  vars_list <- progressr::with_progress({
    p <- progressr::progressor(steps = 7)
    p(message = "Calculating CBD distance");       dist_cbd <- as.numeric(st_distance(area_centroids, cbd_centroid))
    p(message = "Calculating distance to Hospitals"); dist_hosp <- nearest_dist(area_centroids, hosp_sf)
    p(message = "Calculating distance to Parks");     dist_park <- nearest_dist(area_centroids, park_sf)
    p(message = "Calculating distance to Rail stops");  dist_rail <- nearest_dist(area_centroids, rail_sf)
    p(message = "Counting shops within 1 km");     shops_1km <- shop_count(area_centroids, shops_sf, 1000)
    p(message = "Counting shops within 2 km");     shops_2km <- shop_count(area_centroids, shops_sf, 2000)
    p(message = "Counting shops within 5 km");     shops_5km <- shop_count(area_centroids, shops_sf, 5000)

    list(
      dist_cbd_km  = dist_cbd / 1000,
      dist_hosp_km = as.numeric(dist_hosp) / 1000,
      dist_park_km = as.numeric(dist_park) / 1000,
      dist_rail_km = as.numeric(dist_rail) / 1000,
      shops_1km    = shops_1km,
      shops_2km    = shops_2km,
      shops_5km    = shops_5km
    )
  })

  out_sf <- area_sf %>%
    mutate(!!!vars_list)

  return(out_sf)
}


all_sa1 <- add_geo_vars(all_sa1)
all_sa2 <- add_geo_vars(all_sa2)
all_poa <- add_geo_vars(all_poa)
gnaf_sa1_house = add_geo_vars(gnaf_sa1_house)
gnaf_sa1_apt = add_geo_vars(gnaf_sa1_apt)
gnaf_sa2_house = add_geo_vars(gnaf_sa2_house)
gnaf_sa2_apt = add_geo_vars(gnaf_sa2_apt)
gnaf_poa_house = add_geo_vars(gnaf_poa_house)
gnaf_poa_apt = add_geo_vars(gnaf_poa_apt)

transactions_sf$centroid  = transactions_sf$geometry
# 1  Split row indices into 10 roughly equal chunks ------------------
n_chunks <- 10
row_index <- split(
  seq_len(nrow(transactions_sf)),
  cut(seq_len(nrow(transactions_sf)),
      breaks = n_chunks, labels = FALSE)
)

# -------------------------------------------------------------------
# 2  Loop over chunks, run add_geo_vars(), collect results -----------
transactions_geo_list <- vector("list", n_chunks)

for (i in seq_along(row_index)) {
  rows <- row_index[[i]]
  transactions_geo_list[[i]] <- add_geo_vars(transactions_sf[rows, ])
}

# -------------------------------------------------------------------
# 3  Row-bind all chunk results --------------------------------------
transactions_sf <- dplyr::bind_rows(transactions_geo_list)
```

# Merge all

I will merge `transactions` with created variables from `sa1_geo`, `sa2_geo`, and `poa_geo`. They will have `varname_suffix` such as `dist_cbd_km_sa1`.

```{R}
# --------------------------------------------------------------------------
# Merge all geographic levels onto the transaction data (Final Version)
# --------------------------------------------------------------------------
# This version uses the superior function-based approach you designed,
# with one minor refinement to make the helper function fully self-contained.

handlers("txtprogressbar") # A clean progress bar handler

# The variables we want to attach from the area layers
var_interest <- c(
  "dist_cbd_km", "dist_hosp_km", "dist_park_km", "dist_rail_km",
  "shops_1km", "shops_2km", "shops_5km"
)

#' Spatially joins area-level variables to a transaction sf object.
#'
#' @param trans_sf The sf object with transaction points.
#' @param area_sf The sf object with area polygons (e.g., SA1, SA2).
#' @param vars_keep A character vector of columns to keep from `area_sf`.
#' @param suffix A string to append to the new column names.
#' @param code_col The name of the ID column in `area_sf` (e.g., "SA1_CODE21").
#' @return The `trans_sf` object with new columns added.
add_area_vars <- function(trans_sf, area_sf, vars_keep, suffix, code_col) {
  
  # Prepare the area data for joining:
  # 1. Drop the polygon geometry to prevent it from being added to the points.
  # 2. Select only the necessary columns.
  # 3. Add the specified suffix to the variable names.
  area_to_join <- area_sf %>%
    st_drop_geometry() %>%
    select(any_of(c(code_col, vars_keep))) %>%
    rename_with(~ ifelse(.x == code_col, .x, paste0(.x, suffix)))

  # Perform the spatial join. `st_join` will use the geometry from `area_sf`
  # (the original object, before we dropped its geometry in the pipe)
  # to find the matches for `trans_sf`.
  st_join(trans_sf, area_sf %>% select(all_of(code_col))) %>%
    # Now, left-join the prepared attributes using the area code.
    # This is faster and cleaner than joining all attributes spatially.
    left_join(area_to_join, by = code_col)
}

# A slightly improved version of the helper function that is more efficient
add_area_vars_optimized <- function(trans_sf, area_sf, vars_keep, suffix, code_col) {
  
  # Prepare the area data for joining by dropping geometry and renaming
  area_attributes <- area_sf %>%
    st_drop_geometry() %>%
    select(any_of(c(code_col, vars_keep))) %>%
    rename_with(~ ifelse(.x == code_col, .x, paste0(.x, suffix)))

  # Step 1: Perform a lean spatial join to only get the ID of the containing polygon
  trans_with_id <- st_join(trans_sf, area_sf %>% select(all_of(code_col)), left = TRUE)
  
  # Step 2: Perform a fast, non-spatial attribute join using the ID
  left_join(trans_with_id, area_attributes, by = code_col)
}


# --- Execution using the optimized function ---
transactions_final_sf <- progressr::with_progress({
  p <- progressr::progressor(steps = 3)

  # --- SA1: Join all relevant variables ---
  p(message = "Joining SA1 variables...")
  sa1_keep <- setdiff(names(all_sa1), c("geometry", "centroid"))
  transactions_sf <- add_area_vars_optimized(
    transactions_sf, all_sa1, sa1_keep, "_sa1", "SA1_CODE21"
  )

  # --- SA2: Join only the variables of interest ---
  p(message = "Joining SA2 variables...")
  transactions_sf <- add_area_vars_optimized(
    transactions_sf, all_sa2, var_interest, "_sa2", "SA2_CODE21"
  )

  # --- POA: Join the postcode and variables of interest ---
  p(message = "Joining POA variables...")
  transactions_sf <- add_area_vars_optimized(
    transactions_sf, all_poa, var_interest, "_poa", "POA_CODE21"
  )
  
  # Return the final object
  transactions_sf
})

# --- Final Check ---
# The final object is clean, with no extra geometry columns to remove.
glimpse(transactions_final_sf)

```

Now given that `transactions_final_sf` has SA1, SA2 and POA identifiers, we can easily merge them with gnaf-based calculated amenity variables. So, merge `transactions_final_sf` with `gnaf_sa1_apt` and others above. using the identifiers...

```{r}
# --------------------------------------------------------------------------
# Merge G-NAF-based amenity variables (house / apartment) onto transactions
# --------------------------------------------------------------------------

handlers("txtprogressbar")

geo_vars <- c("dist_cbd_km","dist_hosp_km","dist_park_km",
              "dist_rail_km","shops_1km","shops_2km","shops_5km")

# helper → drop geometry & add suffix to the seven variables ---------------
suffix_vars <- function(sf, suffix, code_col) {
  sf %>% 
    st_drop_geometry() %>% 
    rename_with(~ ifelse(.x %in% geo_vars, paste0(.x, suffix), .x),
                all_of(geo_vars)) %>% 
    select(any_of(c(code_col, paste0(geo_vars, suffix))))
}

# build six lean lookup tables ---------------------------------------------
sa1_house_attr <- suffix_vars(gnaf_sa1_house, "_sa1_house", "SA1_CODE21")
sa1_apt_attr   <- suffix_vars(gnaf_sa1_apt,   "_sa1_apt",   "SA1_CODE21")

sa2_house_attr <- suffix_vars(gnaf_sa2_house, "_sa2_house", "SA2_CODE21")
sa2_apt_attr   <- suffix_vars(gnaf_sa2_apt,   "_sa2_apt",   "SA2_CODE21")

poa_house_attr <- suffix_vars(gnaf_poa_house, "_poa_house", "POA_CODE21")
poa_apt_attr   <- suffix_vars(gnaf_poa_apt,   "_poa_apt",   "POA_CODE21")

# join everything onto transactions_final_sf -------------------------------
transactions_final_sf <- progressr::with_progress({
  p <- progressr::progressor(6)
  transactions_final_sf %>% 
    {p("SA1-house"); left_join(., sa1_house_attr, by = "SA1_CODE21")} %>% 
    {p("SA1-apt");   left_join(., sa1_apt_attr,   by = "SA1_CODE21")} %>% 
    {p("SA2-house"); left_join(., sa2_house_attr, by = "SA2_CODE21")} %>% 
    {p("SA2-apt");   left_join(., sa2_apt_attr,   by = "SA2_CODE21")} %>% 
    {p("POA-house"); left_join(., poa_house_attr, by = "POA_CODE21")} %>% 
    {p("POA-apt");   left_join(., poa_apt_attr,   by = "POA_CODE21")}
})

# quick sanity-check
glimpse(transactions_final_sf, width = 80)
```

# Get SEIFA

```{r}
seifa = readxl::read_xlsx('../Data/Inputs/seifa.xlsx')
mean(transactions_final_sf$SA1_CODE21 %in% seifa$SA1)
```

merge them:

```{r}
transactions_final_sf = merge(transactions_final_sf, 
      seifa, 
      by.x = 'SA1_CODE21',
      by.y = 'SA1', 
      all.x = T, 
      all.y = F)
```

Save the file

```{r}
transactions_final_sf %>% select(-centroid) %>%
  st_write('../Data/Intermediate/transactions_with_GNAF_and_amenities_2022_2025_revision.gpkg', 
           delete_dsn = TRUE, quiet = TRUE)
```

# Additional Attributes (2026)

This section creates the regional variables used by
`Regressions_v10.Rmd`. It calculates only the variables needed in the
analysis, reuses the completed Hopkins--Skellam results, and retains the
transaction point geometry in the final GeoPackage.

## Setup

```{r}
library(dplyr)
library(sf)
library(purrr)
library(future)
library(future.apply)
library(progressr)

# Use at most eight workers and leave one logical core free.
n_workers <- min(8, max(1, parallel::detectCores() - 1))
chunk_size <- 100

handlers("txtprogressbar")
sf_use_s2(FALSE)

capital_path <- "../Data/Inputs/Boundaries/GCCSA_2021_AUST_GDA2020.shp"
zoning_path <- "../Data/Inputs/Zoning/EPI_Land_Zoning.shp"
hs_path <- "../Data/Inputs/hopskell_all_mc.rds"
transaction_path <- paste0(
  "../Data/Intermediate/",
  "transactions_with_GNAF_and_amenities_2022_2025_revision.gpkg"
)

geographies <- list(
  SA1 = list(
    path = "../Data/Inputs/Boundaries/sa1_all.gpkg",
    code = "SA1_CODE21",
    prefix = "1"
  ),
  SA2 = list(
    path = "../Data/Inputs/Boundaries/sa2_all.gpkg",
    code = "SA2_CODE21",
    prefix = "1"
  ),
  POA = list(
    path = "../Data/Inputs/Boundaries/poa_all.gpkg",
    code = "POA_CODE21",
    prefix = "2"
  )
)
```

## Shared spatial layers

```{r}
# Greater Sydney boundary
capital_syd <- st_read(capital_path, quiet = TRUE) %>%
  filter(GCC_CODE21 == "1GSYD") %>%
  st_transform(7856) %>%
  st_make_valid()

# R1--R5 are the residential zones used to calculate residential coverage.
# st_make_valid() is applied after projection because projection can introduce
# a small number of invalid geometries even when the source layer is valid.
residential_zones <- st_read(zoning_path, quiet = TRUE) %>%
  filter(SYM_CODE %in% paste0("R", 1:5)) %>%
  select(geometry) %>%
  st_transform(7856) %>%
  st_make_valid()
```

## Functions used for all three geographies

```{r}
# Read one geography, keep the relevant state/postcode observations, and
# retain only its identifier and geometry.
load_geography <- function(path, code_col, prefix) {
  geography <- st_read(path, quiet = TRUE)
  keep <- startsWith(as.character(geography[[code_col]]), prefix)
  geography <- geography[keep, code_col]

  geography %>%
    st_transform(7856) %>%
    st_make_valid()
}

# Residential area within every boundary. This replaces the old row-by-row
# st_difference() calculation. Direct intersection correctly returns zero
# residential area when no residential zoning overlaps a boundary.
residential_coverage <- function(geography, code_col, residential_zones) {
  boundaries <- st_sf(
    code = as.character(geography[[code_col]]),
    geometry = st_geometry(geography)
  )

  residential <- st_sf(geometry = st_geometry(residential_zones))

  pieces <- suppressWarnings(st_intersection(boundaries, residential))

  if (nrow(pieces) == 0) {
    return(rep(0, nrow(geography)))
  }

  coverage <- pieces %>%
    group_by(code) %>%
    summarise(.groups = "drop") %>%
    mutate(residential_area = as.numeric(st_area(geometry))) %>%
    st_drop_geometry()

  tibble(code = as.character(geography[[code_col]])) %>%
    left_join(coverage, by = "code") %>%
    mutate(residential_area = coalesce(residential_area, 0)) %>%
    pull(residential_area)
}

# Circularity for one polygon. The fitted circle uses the polygon centroid and
# the root-mean-square distance from that centroid to the boundary vertices.
# The numerator is the symmetric-difference area:
# area(polygon) + area(circle) - 2 * area(polygon intersect circle).
circularity_one <- function(geometry) {
  tryCatch({
    polygon_area <- as.numeric(sf::st_area(geometry))
    coordinates <- sf::st_coordinates(geometry)[, c("X", "Y"), drop = FALSE]
    centre <- suppressWarnings(sf::st_centroid(geometry))
    centre_xy <- sf::st_coordinates(centre)[1, c("X", "Y")]

    radius <- sqrt(mean(
      (coordinates[, "X"] - centre_xy[["X"]])^2 +
        (coordinates[, "Y"] - centre_xy[["Y"]])^2
    ))

    circle <- sf::st_buffer(centre, dist = radius)
    overlap <- suppressWarnings(sf::st_intersection(geometry, circle))
    overlap_area <- if (length(overlap) == 0 || all(sf::st_is_empty(overlap))) {
      0
    } else {
      sum(as.numeric(sf::st_area(overlap)), na.rm = TRUE)
    }

    circle_area <- as.numeric(sf::st_area(circle))
    symmetric_difference <- polygon_area + circle_area - 2 * overlap_area

    max(symmetric_difference, 0) / polygon_area
  }, error = function(e) NA_real_)
}

# Each worker receives only a small geometry chunk and returns a numeric vector.
# This avoids copying the zoning and transaction data to every worker.
circularity_chunk <- function(geometry_chunk) {
  vapply(
    seq_along(geometry_chunk),
    function(i) circularity_one(geometry_chunk[i]),
    numeric(1)
  )
}

parallel_circularity <- function(geometry, chunk_size) {
  row_groups <- split(
    seq_along(geometry),
    ceiling(seq_along(geometry) / chunk_size)
  )
  geometry_chunks <- lapply(row_groups, function(rows) geometry[rows])

  with_progress({
    progress <- progressor(steps = length(geometry_chunks))

    values <- future_lapply(
      geometry_chunks,
      function(one_chunk) {
        answer <- circularity_chunk(one_chunk)
        progress()
        answer
      },
      future.scheduling = Inf,
      future.seed = TRUE
    )
  }, enable = TRUE)

  unlist(values, use.names = FALSE)
}

process_geography <- function(details, label, residential_zones, capital_syd) {
  geography <- load_geography(
    details$path,
    details$code,
    details$prefix
  )

  message("\n", label, ": ", nrow(geography), " boundaries")

  message(label, " — 1/3: area and Greater Sydney indicator")
  area_sa <- as.numeric(st_area(geography))
  sydney <- lengths(st_intersects(geography, capital_syd)) > 0

  message(label, " — 2/3: residential coverage")
  area_resid <- residential_coverage(
    geography,
    details$code,
    residential_zones
  )
  shr_resid <- ifelse(area_sa > 0, area_resid / area_sa, NA_real_)
  shr_resid <- pmin(pmax(shr_resid, 0), 1)

  message(
    label, " — 3/3: circularity on ", n_workers,
    " workers (progress is by ", chunk_size, "-boundary chunk)"
  )
  var_circ <- parallel_circularity(
    st_geometry(geography),
    chunk_size
  )

  tibble(
    !!details$code := as.character(geography[[details$code]]),
    sydney = sydney,
    area_sa = area_sa,
    var_circ = var_circ,
    shr_resid = shr_resid
  )
}
```

## Calculate SA1, SA2 and POA attributes

The zoning overlay is vectorised. Circularity is the only calculation run in
parallel because it requires a separate fitted circle for each boundary.

```{r}
plan(multisession, workers = n_workers)

regional_attributes <- imap(
  geographies,
  ~ process_geography(
    details = .x,
    label = .y,
    residential_zones = residential_zones,
    capital_syd = capital_syd
  )
)

plan(sequential)
```

## Add the completed Hopkins--Skellam results

The Hopkins--Skellam simulations are not rerun here. The completed results in
`hopskell_all_mc.rds` are joined to the newly calculated regional attributes.

```{r}
hs <- readRDS(hs_path) %>%
  transmute(
    zone_type,
    zone_code = as.character(zone_code),
    HS_stat_A = as.numeric(H),
    HS_test_pval = as.numeric(p_value)
  )

add_hs <- function(attributes, label, code_col) {
  hs_level <- hs %>%
    filter(zone_type == label) %>%
    select(-zone_type)

  left_join(
    attributes,
    hs_level,
    by = setNames("zone_code", code_col)
  )
}

sa1 <- add_hs(regional_attributes$SA1, "SA1", "SA1_CODE21")
sa2 <- add_hs(regional_attributes$SA2, "SA2", "SA2_CODE21")
poa <- add_hs(regional_attributes$POA, "POA", "POA_CODE21")

saveRDS(sa1, "../Data/Intermediate/sa1_nsw.rds")
saveRDS(sa2, "../Data/Intermediate/sa2_nsw.rds")
saveRDS(poa, "../Data/Intermediate/poa_nsw.rds")
```

## Add the regional attributes to the transaction data

The point geometry is retained so that `Regressions_v10.Rmd` can generate the
transaction map from the same analysis file.

```{r}
transactions <- st_read(transaction_path, quiet = TRUE) %>%
  filter(STE_CODE21_sa1 == "1")

regional_for_join <- function(attributes, code_col, prefix) {
  attributes %>%
    select(
      all_of(code_col),
      sydney, area_sa, var_circ, shr_resid,
      HS_stat_A, HS_test_pval
    ) %>%
    rename_with(
      ~ paste0(prefix, "_", .x),
      -all_of(code_col)
    )
}

transactions_updated <- transactions %>%
  left_join(
    regional_for_join(sa1, "SA1_CODE21", "sa1"),
    by = "SA1_CODE21"
  ) %>%
  left_join(
    regional_for_join(sa2, "SA2_CODE21", "sa2"),
    by = "SA2_CODE21"
  ) %>%
  left_join(
    regional_for_join(poa, "POA_CODE21", "poa"),
    by = "POA_CODE21"
  )

output_path <- paste0(
  "../Data/Analysis/",
  "transactions_with_GNAF_and_amenities_2022_2025_nsw_updated_20260726.gpkg"
)

st_write(
  transactions_updated,
  output_path,
  delete_dsn = TRUE,
  quiet = TRUE
)
```

## Simple checks

```{r}
required_regional_columns <- unlist(lapply(
  c("sa1", "sa2", "poa"),
  function(level) {
    paste0(
      level, "_",
      c("sydney", "area_sa", "var_circ", "shr_resid", "HS_stat_A")
    )
  }
))

cat("Input transaction rows: ", nrow(transactions), "\n")
cat("Output transaction rows:", nrow(transactions_updated), "\n")
cat(
  "Missing required columns:",
  paste(setdiff(required_regional_columns, names(transactions_updated)),
        collapse = ", "),
  "\n"
)
cat(
  "Geometry retained:",
  inherits(transactions_updated, "sf"),
  "—", paste(unique(as.character(st_geometry_type(transactions_updated))),
             collapse = ", "),
  "\n"
)

transactions_updated %>%
  st_drop_geometry() %>%
  summarise(
    sa1_share_min = min(sa1_shr_resid, na.rm = TRUE),
    sa1_share_max = max(sa1_shr_resid, na.rm = TRUE),
    sa2_share_min = min(sa2_shr_resid, na.rm = TRUE),
    sa2_share_max = max(sa2_shr_resid, na.rm = TRUE),
    poa_share_min = min(poa_shr_resid, na.rm = TRUE),
    poa_share_max = max(poa_shr_resid, na.rm = TRUE)
  )
```
