---
title: "Regression and publication outputs — v10"
output:
  html_notebook:
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: inline
---

# Purpose

This notebook runs the complete analysis and creates the publication-ready
tables and figures in `Outputs`. The code uses one set of
plain labels throughout:

\[
S_i=\text{exact-coordinate spatial variable}, \qquad
\widetilde S_i^g=\text{centroid proxy}, \qquad
e_i^g=\widetilde S_i^g-S_i.
\]

The regression comparison is

\[
\Delta^g=widehat\beta_{\text{centroid},g}
-\widehat\beta_{\text{exact}}.
\]

Thus, \(\Delta=0\) means that centroid substitution does not change the
coefficient. Raw coefficient gaps remain the inferential object. Relative
coefficient changes are also reported for the two primary variables:

\[
P^g=100\left(
\frac{\widehat\beta_{\text{centroid},g}}
{\widehat\beta_{\text{exact}}}-1
\right).
\]

```{r setup}
#| message: false
#| warning: false

packages_used <- c(
  "data.table", "dplyr", "sf", "fixest", "modelsummary",
  "ggplot2", "sandwich"
)

suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(sf)
  library(fixest)
  library(modelsummary)
  library(ggplot2)
})

options(
  modelsummary_format_numeric_latex = "plain",
  scipen = 999
)

REGRESSION_DIR <- if (file.exists("Regressions_v10.Rmd")) {
  normalizePath(".")
} else {
  normalizePath("Code")
}
PROJECT_ROOT <- normalizePath(file.path(REGRESSION_DIR, ".."))
DATA_FILE <- file.path(
  PROJECT_ROOT, "Data", "Analysis",
  "transactions_with_GNAF_and_amenities_2022_2025_nsw_updated_20260726.gpkg"
)

BOOTSTRAP_REPS <- 999L
BOOTSTRAP_SEED <- 20260714L

OUTPUT_DIR <- file.path(PROJECT_ROOT, "Outputs")
TABLE_DIR <- file.path(OUTPUT_DIR, "tables")
FIGURE_DIR <- file.path(OUTPUT_DIR, "figures")
DATA_DIR <- file.path(OUTPUT_DIR, "data")

for (path in c(OUTPUT_DIR, TABLE_DIR, FIGURE_DIR, DATA_DIR)) {
  dir.create(path, recursive = TRUE, showWarnings = FALSE)
}

# Start each run with a clean set of generated tables, figures, and CSV files.
for (path in c(TABLE_DIR, FIGURE_DIR, DATA_DIR)) {
  unlink(list.files(path, full.names = TRUE))
}
```

# Configuration

```{r specification-dictionaries}
outcomes <- data.table(
  outcome_key = c("cbd", "shops1", "shops2", "shops5", "hospital", "park", "rail"),
  outcome = c(
    "CBD distance", "Shops within 1 km", "Shops within 2 km",
    "Shops within 5 km", "Hospital distance", "Park distance",
    "Rail distance"
  ),
  stem = c(
    "dist_cbd_km", "shops_1km", "shops_2km", "shops_5km",
    "dist_hosp_km", "dist_park_km", "dist_rail_km"
  ),
  family = c("primary", "primary", "primary", "primary",
             "additional", "additional", "additional"),
  outcome_order = seq_len(7L)
)

geographies <- data.table(
  geo_key = c("sa1", "sa2", "poa"),
  geography = c("SA1", "SA2", "Postcode"),
  suffix = c("sa1", "sa2", "poa"),
  cluster_var = c("SA1_CODE21", "SA2_CODE21", "POA_CODE21"),
  id_var = c("SA1_CODE21", "SA2_CODE21", "POA_CODE21"),
  attribute_prefix = c("sa1", "sa2", "poa"),
  geo_order = seq_len(3L)
)

property_types <- data.table(
  property_key = c("house", "apartment"),
  property_type = c("House", "Apartment"),
  house_value = c(TRUE, FALSE),
  property_order = seq_len(2L)
)

fixed_effects <- data.table(
  fe_key = c("none", "sa3", "sa4"),
  fixed_effects = c("No FE", "SA3", "SA4"),
  fe_rhs = c("", "SA3_CODE21_sa1 + timeFE", "SA4_CODE21_sa1 + timeFE"),
  fe_order = seq_len(3L)
)

sample_definitions <- data.table(
  sample_key = c(
    "full", "sydney", "high_clustering", "high_irregularity",
    "low_residential", "large_area"
  ),
  sample = c(
    "Full sample", "Greater Sydney", "High clustering",
    "Most irregular", "Low residential share", "Largest area"
  ),
  sample_order = seq_len(6L)
)

cross_join <- function(x, y) {
  data.table::as.data.table(merge(
    as.data.frame(x), as.data.frame(y), by = NULL
  ))
}

controls_for <- function(stem) {
  common <- c("splines::ns(Area, df = 3)", "splines::ns(ier, df = 3)")
  extras <- switch(
    stem,
    dist_cbd_km = c(
      "log(dist_hosp_km * 1000 + 1)",
      "log(dist_park_km * 1000 + 1)",
      "log(dist_rail_km * 1000 + 1)",
      "log(shops_1km + 1)"
    ),
    shops_1km = c(
      "log(dist_hosp_km * 1000 + 1)",
      "log(dist_park_km * 1000 + 1)",
      "log(dist_rail_km * 1000 + 1)",
      "log(dist_cbd_km * 1000 + 1)"
    ),
    shops_2km = c(
      "log(dist_hosp_km * 1000 + 1)",
      "log(dist_park_km * 1000 + 1)",
      "log(dist_rail_km * 1000 + 1)",
      "log(dist_cbd_km * 1000 + 1)"
    ),
    shops_5km = c(
      "log(dist_hosp_km * 1000 + 1)",
      "log(dist_park_km * 1000 + 1)",
      "log(dist_rail_km * 1000 + 1)",
      "log(dist_cbd_km * 1000 + 1)"
    ),
    dist_hosp_km = c(
      "log(dist_park_km * 1000 + 1)",
      "log(dist_rail_km * 1000 + 1)",
      "log(dist_cbd_km * 1000 + 1)",
      "log(shops_1km + 1)"
    ),
    dist_park_km = c(
      "log(dist_hosp_km * 1000 + 1)",
      "log(dist_rail_km * 1000 + 1)",
      "log(dist_cbd_km * 1000 + 1)",
      "log(shops_1km + 1)"
    ),
    dist_rail_km = c(
      "log(dist_hosp_km * 1000 + 1)",
      "log(dist_park_km * 1000 + 1)",
      "log(dist_cbd_km * 1000 + 1)",
      "log(shops_1km + 1)"
    )
  )
  c(common, extras)
}

transform_spatial <- function(x, stem) {
  if (startsWith(stem, "dist_")) log(x * 1000 + 1) else log(x + 1)
}
```

# Spatial-variable constructions

For amenity type \(k\), exact and centroid-based nearest distances are

\[
D_{ik}=\min_j d(p_i,a_{jk}), \qquad
\widetilde D_{ik}^g=\min_j d(c_{g(i)},a_{jk}),
\]

where \(p_i\) is the exact property coordinate, \(c_{g(i)}\) is the centroid
of its released geography, and \(a_{jk}\) is an amenity location.

For shops inside radius \(r\),

\[
C_i(r)=\sum_j \mathbf 1\{d(p_i,a_j)\le r\}, \qquad
\widetilde C_i^g(r)=
\sum_j \mathbf 1\{d(c_{g(i)},a_j)\le r\}.
\]

The one-, two-, and five-kilometre comparisons show whether centroid
sensitivity changes as the constructed variable operates over a larger
spatial scale.

# Load and clean data once

```{r load-and-clean-data}

base_columns <- c(
  "SA1_CODE21", "SA2_CODE21", "POA_CODE21", "SA3_CODE21_sa1",
  "SA4_CODE21_sa1", "Property.unit.number", "Purchase.price", "Area",
  "Contract.date", "Settlement.date", "irsd", "irsad", "ier",
  paste0(rep(c("sa1", "sa2", "poa"), each = 5L), "_",
         rep(c("sydney", "area_sa", "var_circ", "shr_resid", "HS_stat_A"), 3L))
)

identifier_columns <- c(
  "address", "Dealing.number", "Download.date...time",
  "ADDRESS_DETAIL_PID", "Property.ID", "Sale.counter"
)

all_stems <- c(
  "dist_cbd_km", "dist_hosp_km", "dist_park_km", "dist_rail_km",
  "shops_1km", "shops_2km", "shops_5km"
)
spatial_columns <- unique(c(
  all_stems,
  unlist(lapply(all_stems, function(stem) paste0(stem, "_", c("sa1", "sa2", "poa")))),
  unlist(lapply(all_stems, function(stem) {
    paste0(stem, "_", rep(c("sa1", "sa2", "poa"), each = 2L),
           "_", rep(c("house", "apt"), 3L))
  }))
))

layer_name <- sf::st_layers(DATA_FILE)$name[[1L]]
quote_identifier <- function(x) paste0('"', gsub('"', '""', x, fixed = TRUE), '"')

requested_columns <- unique(c(identifier_columns, base_columns, spatial_columns))

select_sql <- paste(quote_identifier(requested_columns), collapse = ", ")
analysis_query <- sprintf(
  "SELECT %s FROM %s",
  select_sql,
  quote_identifier(layer_name)
)

transactions_raw <- sf::st_read(DATA_FILE, query = analysis_query, quiet = TRUE)
if (inherits(transactions_raw, "sf")) {
  transactions_raw <- sf::st_drop_geometry(transactions_raw)
}
transactions <- data.table::as.data.table(transactions_raw)
rm(transactions_raw)

input_rows <- nrow(transactions)
transactions[, sale_id := paste(Dealing.number, address, sep = "|")]
input_sale_ids <- uniqueN(transactions$sale_id)

eligible_input_rows <- transactions[
  Purchase.price >= 200000 & Purchase.price < 5000000 &
    Area < 10000 & !is.na(Purchase.price) & !is.na(Area),
  .N
]

# Keep the most recently downloaded version of each dealing-address sale.
# All rows from that download are retained here so that non-unique G-NAF
# matches can be identified rather than resolved arbitrarily.
transactions[, latest_download := max(Download.date...time), by = sale_id]
transactions <- transactions[Download.date...time == latest_download]

transactions <- transactions[
  Purchase.price >= 200000 & Purchase.price < 5000000 &
    Area < 10000 & !is.na(Purchase.price) & !is.na(Area)
]

eligible_latest_rows <- nrow(transactions)
eligible_sale_ids <- uniqueN(transactions$sale_id)

# Rows that differ only in source identifiers have identical inputs for every
# analysis in this notebook and can be collapsed safely. If a sale retains
# more than one distinct set of model inputs, its exact-coordinate benchmark
# is ambiguous and the sale is excluded from the primary validation sample.
analysis_columns <- unique(c(base_columns, spatial_columns))
transactions_distinct <- unique(
  transactions,
  by = c("sale_id", analysis_columns)
)

candidate_counts <- transactions_distinct[, .(candidate_rows = .N), by = sale_id]
ambiguous_sale_ids <- candidate_counts[candidate_rows > 1L, sale_id]

ambiguous_candidates <- transactions_distinct[sale_id %chin% ambiguous_sale_ids]
ambiguous_match_summary <- ambiguous_candidates[, {
  exact_counts <- vapply(.SD, uniqueN, integer(1L))
  list(
    property_type = ifelse(Property.unit.number[[1L]] == "", "House", "Apartment"),
    conflict_type = ifelse(
      any(exact_counts > 1L),
      "Conflicting exact spatial variables",
      "Other conflicting model inputs"
    ),
    conflicting_spatial_variables = paste(
      names(exact_counts)[exact_counts > 1L],
      collapse = ", "
    ),
    candidate_rows = .N
  )
}, by = sale_id, .SDcols = all_stems]

clean_sale_ids <- candidate_counts[candidate_rows == 1L, sale_id]
transactions <- transactions_distinct[sale_id %chin% clean_sale_ids]

transactions[, `:=`(
  ier = suppressWarnings(as.numeric(ier)),
  irsd = suppressWarnings(as.numeric(irsd)),
  irsad = suppressWarnings(as.numeric(irsad)),
  house = Property.unit.number == "",
  log_price = log(Purchase.price),
  timeFE = format(as.IDate(Contract.date), "%Y-%m")
)]

transactions[, property_type := fifelse(house, "House", "Apartment")]
setkey(transactions, house)

sample_counts <- transactions[, .N, by = property_type][order(property_type)]
ambiguous_counts <- ambiguous_match_summary[, .N, by = .(property_type, conflict_type)]

eligible_property_counts <- unique(
  transactions_distinct[, .(
    sale_id,
    property_type = fifelse(Property.unit.number == "", "House", "Apartment")
  )],
  by = "sale_id"
)[, .(eligible_sales = .N), by = property_type]

ambiguity_by_property <- merge(
  eligible_property_counts,
  ambiguous_match_summary[, .(excluded_sales = .N), by = property_type],
  by = "property_type",
  all = TRUE
)
ambiguity_by_property <- merge(
  ambiguity_by_property,
  sample_counts[, .(property_type, final_sales = N)],
  by = "property_type",
  all = TRUE
)
ambiguity_by_property[, excluded_percent := 100 * excluded_sales / eligible_sales]

sample_cleaning_flow <- data.table(
  stage = c(
    "Rows in matched analysis file",
    "Distinct dealing-address sale identifiers",
    "Rows satisfying the original price and area restrictions",
    "Latest eligible dealing-address sales",
    "Excluded sales with conflicting model inputs",
    "Final unambiguous validation sample",
    "Final houses",
    "Final apartments"
  ),
  unit = c("Rows", "Sales", "Rows", "Sales", "Sales", "Sales", "Sales", "Sales"),
  count = c(
    input_rows,
    input_sale_ids,
    eligible_input_rows,
    eligible_sale_ids,
    length(ambiguous_sale_ids),
    nrow(transactions),
    transactions[house == TRUE, .N],
    transactions[house == FALSE, .N]
  )
)

sample_counts
sample_cleaning_flow
ambiguous_counts
ambiguity_by_property
```

# Descriptive statistics

```{r descriptive-statistics}
descriptive_dictionary <- data.table(
  variable = c(
    "Purchase.price", "Area", "irsad", "ier", "dist_cbd_km",
    "dist_hosp_km", "dist_park_km", "dist_rail_km", "shops_1km"
  ),
  label = c(
    "Price (AUD)", "Area (sqm)", "IRSAD", "IER", "CBD distance (km)",
    "Hospital distance (km)", "Park distance (km)",
    "Rail distance (km)", "Shops within 1 km"
  ),
  source = c(
    "NSW Valuer General", "NSW Valuer General", "ABS", "ABS",
    "G-NAF", "OpenStreetMap", "OpenStreetMap", "OpenStreetMap", "G-NAF"
  )
)

descriptive_long <- rbindlist(lapply(seq_len(nrow(descriptive_dictionary)), function(i) {
  var <- descriptive_dictionary$variable[[i]]
  transactions[, .(
    mean = as.numeric(mean(get(var), na.rm = TRUE)),
    median = as.numeric(median(get(var), na.rm = TRUE)),
    sd = as.numeric(sd(get(var), na.rm = TRUE)),
    N = sum(!is.na(get(var)))
  ), by = property_type][, `:=`(
    variable = var,
    label = descriptive_dictionary$label[[i]],
    source = descriptive_dictionary$source[[i]]
  )]
}), fill = TRUE)

descriptive_table <- dcast(
  descriptive_long,
  label + source ~ property_type,
  value.var = c("mean", "median", "sd", "N")
)
descriptive_table[, display_order := match(label, descriptive_dictionary$label)]
setorder(descriptive_table, display_order)
```

# Geography-based sample definitions

```{r sample-membership}
build_sample_ids <- function(data, geo_row) {
  id_var <- geo_row$id_var
  prefix <- geo_row$attribute_prefix
  attributes <- c(
    paste0(prefix, "_sydney"),
    paste0(prefix, "_area_sa"),
    paste0(prefix, "_var_circ"),
    paste0(prefix, "_shr_resid"),
    paste0(prefix, "_HS_stat_A")
  )
  unit_data <- unique(data[, c(id_var, attributes), with = FALSE], by = id_var)

  q25_hs <- quantile(unit_data[[paste0(prefix, "_HS_stat_A")]], 0.25, na.rm = TRUE)
  q75_irregular <- quantile(unit_data[[paste0(prefix, "_var_circ")]], 0.75, na.rm = TRUE)
  q25_residential <- quantile(unit_data[[paste0(prefix, "_shr_resid")]], 0.25, na.rm = TRUE)
  q75_area <- quantile(unit_data[[paste0(prefix, "_area_sa")]], 0.75, na.rm = TRUE)

  list(
    full = unit_data[[id_var]],
    sydney = unit_data[get(paste0(prefix, "_sydney")) == TRUE][[id_var]],
    high_clustering = unit_data[
      get(paste0(prefix, "_HS_stat_A")) <= q25_hs
    ][[id_var]],
    high_irregularity = unit_data[
      get(paste0(prefix, "_var_circ")) >= q75_irregular
    ][[id_var]],
    low_residential = unit_data[
      get(paste0(prefix, "_shr_resid")) <= q25_residential
    ][[id_var]],
    large_area = unit_data[
      get(paste0(prefix, "_area_sa")) >= q75_area
    ][[id_var]]
  )
}

sample_ids <- setNames(vector("list", nrow(geographies)), geographies$geo_key)
for (i in seq_len(nrow(geographies))) {
  sample_ids[[geographies$geo_key[[i]]]] <- build_sample_ids(
    transactions, geographies[i]
  )
}
```

# Generic paired-model engine

The exact and centroid regressions use the same observations, controls, and
fixed effects. Their coefficients are therefore paired rather than
independent. For bootstrap draw \(b\), the same cluster multiplier
\(\xi_g^{(b)}\in\{-1,1\}\) is applied to both coefficient influence
contributions:

\[
\widehat\beta_{exact}^{*(b)}
=\widehat\beta_{exact}
+\sum_g\xi_g^{(b)}\widehat\psi_{exact,g},
\]

\[
\widehat\beta_{centroid}^{*(b)}
=\widehat\beta_{centroid}
+\sum_g\xi_g^{(b)}\widehat\psi_{centroid,g}.
\]

The paired bootstrap draw is

\[
\widehat\Delta^{*(b)}
=\widehat\beta_{centroid}^{*(b)}
-\widehat\beta_{exact}^{*(b)}.
\]

Using the same multiplier preserves covariance between the two estimates.
The bootstrap clusters at the geography supplying the centroid because the
proxy error is shared within that geography. This is an influence-function
multiplier bootstrap; models are not re-estimated 999 times. The resulting
intervals allow arbitrary dependence within the released-geography cluster
and assume independence across those clusters after conditioning on the
controls and fixed effects. They are therefore one-way geography-clustered
intervals, not multiway-cluster counterparts to the coefficient-specific
standard errors reported for the fixed-effect models.

```{r paired-model-functions}
make_formula <- function(outcome_name, controls, fe_rhs = "") {
  rhs <- paste(c(outcome_name, controls), collapse = " + ")
  if (nzchar(fe_rhs)) {
    as.formula(paste("log_price ~", rhs, "|", fe_rhs))
  } else {
    as.formula(paste("log_price ~", rhs))
  }
}

vcov_for <- function(fe_rhs) {
  if (!nzchar(fe_rhs)) "iid" else as.formula(paste0("~", fe_rhs))
}

prepare_pair_data <- function(spec, sample_key = "full", corrected = FALSE) {
  stem <- spec$stem
  proxy_var <- paste0(stem, "_", spec$suffix)
  corrected_var <- paste0(
    proxy_var, "_", if (spec$property_key == "house") "house" else "apt"
  )

  data <- transactions[house == spec$house_value]
  if (sample_key != "full") {
    ids <- sample_ids[[spec$geo_key]][[sample_key]]
    data <- data[get(spec$id_var) %chin% ids]
  }

  keep <- unique(c(
    "log_price", "Area", "ier", "timeFE", "SA3_CODE21_sa1",
    "SA4_CODE21_sa1", spec$cluster_var, stem, proxy_var,
    if (corrected) corrected_var else character(),
    "dist_cbd_km", "dist_hosp_km", "dist_park_km", "dist_rail_km",
    "shops_1km"
  ))
  data <- copy(data[, ..keep])
  data[, `.exact` := transform_spatial(get(stem), stem)]
  data[, `.proxy` := transform_spatial(get(proxy_var), stem)]
  if (corrected) {
    data[, `.weighted` := transform_spatial(get(corrected_var), stem)]
  }

  controls <- controls_for(stem)
  formulas <- list(
    exact = make_formula(".exact", controls, spec$fe_rhs),
    proxy = make_formula(".proxy", controls, spec$fe_rhs)
  )
  if (corrected) {
    formulas$weighted <- make_formula(".weighted", controls, spec$fe_rhs)
  }

  needed <- unique(c(
    unlist(lapply(formulas, all.vars)), spec$cluster_var,
    stem, proxy_var, if (corrected) corrected_var else character()
  ))
  data <- data[complete.cases(data[, ..needed])]
  if (nrow(data) < 30L) stop("Fewer than 30 common complete observations")

  list(
    data = data,
    formulas = formulas,
    vcov = vcov_for(spec$fe_rhs),
    stem = stem,
    proxy_var = proxy_var,
    corrected_var = corrected_var
  )
}

coefficient_row <- function(model, term) {
  # Some small two-way-cluster subsamples require fixest's standard positive-
  # semidefinite VCOV repair; the repaired estimates remain available here.
  estimates <- suppressWarnings(
    modelsummary::get_estimates(model, conf_level = 0.95)
  )
  row <- estimates[estimates$term == term, , drop = FALSE]
  if (nrow(row) != 1L) stop("Focal coefficient not found: ", term)
  c(
    estimate = row$estimate[[1L]],
    std.error = row$std.error[[1L]],
    p.value = row$p.value[[1L]]
  )
}

get_focal_influence <- function(model, coefficient) {
  score <- sandwich::estfun(model)
  bread <- sandwich::bread(model)
  if (is.null(colnames(score))) colnames(score) <- names(coef(model))
  if (is.null(colnames(bread))) {
    colnames(bread) <- rownames(bread) <- names(coef(model))
  }
  if (!coefficient %in% colnames(score) ||
      !coefficient %in% colnames(bread)) {
    stop("Focal term is absent from influence matrices")
  }
  drop(score %*% bread[, coefficient, drop = FALSE]) / nobs(model)
}

paired_multiplier <- function(model_exact, model_proxy, data, cluster_var, B, seed,
                              keep_draws = FALSE) {
  influence_exact <- get_focal_influence(model_exact, ".exact")
  influence_proxy <- get_focal_influence(model_proxy, ".proxy")
  observations <- fixest::obs(model_exact)

  cluster <- as.character(data[[cluster_var]][observations])
  cluster_exact <- rowsum(influence_exact, cluster, reorder = FALSE)[, 1L]
  cluster_proxy <- rowsum(influence_proxy, cluster, reorder = FALSE)[, 1L]
  cluster_exact <- cluster_exact - mean(cluster_exact)
  cluster_proxy <- cluster_proxy - mean(cluster_proxy)
  G <- length(cluster_exact)

  set.seed(seed)
  perturb_exact <- perturb_proxy <- numeric(B)
  batch_size <- 100L
  for (first in seq.int(1L, B, by = batch_size)) {
    index <- first:min(first + batch_size - 1L, B)
    weights <- matrix(
      sample(c(-1, 1), G * length(index), replace = TRUE),
      nrow = G,
      ncol = length(index)
    )
    finite_sample <- sqrt(G / (G - 1))
    perturb_exact[index] <- as.vector(
      crossprod(cluster_exact, weights)
    ) * finite_sample
    perturb_proxy[index] <- as.vector(
      crossprod(cluster_proxy, weights)
    ) * finite_sample
  }

  beta_exact <- unname(coef(model_exact)[[".exact"]])
  beta_proxy <- unname(coef(model_proxy)[[".proxy"]])
  delta <- beta_proxy - beta_exact
  draw_exact <- beta_exact + perturb_exact
  draw_proxy <- beta_proxy + perturb_proxy
  draw_delta <- draw_proxy - draw_exact
  perturb_delta <- draw_delta - delta

  summary <- data.table(
    bootstrap_se = sd(perturb_delta),
    ci_low = unname(quantile(draw_delta, 0.025)),
    ci_high = unname(quantile(draw_delta, 0.975)),
    p_value_delta = (1 + sum(abs(perturb_delta) >= abs(delta))) / (B + 1),
    clusters = G
  )
  draws <- NULL
  if (keep_draws) {
    draws <- data.table(
      draw = seq_len(B),
      beta_exact = draw_exact,
      beta_centroid = draw_proxy,
      delta = draw_delta
    )
  }
  list(summary = summary, draws = draws)
}

absorption_r2 <- function(data, fe_rhs) {
  if (!nzchar(fe_rhs)) return(NA_real_)
  formula <- as.formula(paste(".proxy ~ 1 |", fe_rhs))
  model <- feols(formula, data = data, vcov = "iid", notes = FALSE, warn = FALSE)
  as.numeric(fixest::fitstat(model, "r2")$r2)
}

fit_pair <- function(spec, sample_key = "full", bootstrap = FALSE,
                     keep_draws = FALSE, seed = BOOTSTRAP_SEED) {
  prepared <- prepare_pair_data(spec, sample_key = sample_key)
  data <- prepared$data
  if (uniqueN(data$.proxy) < 2L) {
    return(list(
      summary = data.table(
        status = "not estimable: no focal variation",
        N = nrow(data)
      ),
      draws = NULL
    ))
  }

  model_exact <- feols(
    prepared$formulas$exact, data = data, vcov = prepared$vcov,
    notes = FALSE, warn = FALSE
  )
  model_proxy <- feols(
    prepared$formulas$proxy, data = data, vcov = prepared$vcov,
    notes = FALSE, warn = FALSE
  )
  beta_exact <- coefficient_row(model_exact, ".exact")
  beta_proxy <- coefficient_row(model_proxy, ".proxy")

  result <- data.table(
    status = "estimated",
    benchmark = beta_exact[["estimate"]],
    benchmark_se = beta_exact[["std.error"]],
    benchmark_p = beta_exact[["p.value"]],
    centroid = beta_proxy[["estimate"]],
    centroid_se = beta_proxy[["std.error"]],
    centroid_p = beta_proxy[["p.value"]],
    delta = beta_proxy[["estimate"]] - beta_exact[["estimate"]],
    relative_change_percent =
      100 * (beta_proxy[["estimate"]] / beta_exact[["estimate"]] - 1),
    se_difference =
      beta_proxy[["std.error"]] - beta_exact[["std.error"]],
    absorption_r2 = absorption_r2(data, spec$fe_rhs),
    N = nobs(model_exact)
  )

  draws <- NULL
  if (bootstrap) {
    boot <- paired_multiplier(
      model_exact, model_proxy, data, spec$cluster_var,
      B = BOOTSTRAP_REPS, seed = seed, keep_draws = keep_draws
    )
    result <- cbind(result, boot$summary)
    draws <- boot$draws
  }
  list(summary = result, draws = draws)
}

spec_metadata <- function(spec, sample_key = "full") {
  data.table(
    outcome_key = spec$outcome_key,
    outcome = spec$outcome,
    family = spec$family,
    outcome_order = spec$outcome_order,
    geo_key = spec$geo_key,
    geography = spec$geography,
    geo_order = spec$geo_order,
    property_key = spec$property_key,
    property_type = spec$property_type,
    property_order = spec$property_order,
    fe_key = spec$fe_key,
    fixed_effects = spec$fixed_effects,
    fe_order = spec$fe_order,
    sample_key = sample_key
  )
}

fit_specification <- function(spec, sample_key = "full", bootstrap = FALSE,
                              keep_draws = FALSE, seed = BOOTSTRAP_SEED) {
  metadata <- spec_metadata(spec, sample_key)
  fit <- fit_pair(
    spec, sample_key = sample_key, bootstrap = bootstrap,
    keep_draws = keep_draws, seed = seed
  )
  list(
    summary = cbind(metadata, fit$summary),
    draws = if (is.null(fit$draws)) NULL else cbind(metadata, fit$draws)
  )
}
```

# Full-sample coefficient comparisons and paired inference

```{r full-sample-paired-results}
full_specs <- cross_join(
  cross_join(cross_join(outcomes, geographies), property_types),
  fixed_effects
)
setorder(full_specs, outcome_order, geo_order, property_order, fe_order)
full_specs[, spec_id := .I]

representative_distribution <- function(spec) {
  (spec$outcome_key == "cbd" & spec$geo_key == "sa1" &
     spec$property_key == "house" & spec$fe_key == "none") ||
  (spec$outcome_key == "cbd" & spec$geo_key == "poa" &
     spec$property_key == "apartment" & spec$fe_key == "sa4") ||
  (spec$outcome_key == "shops1" & spec$geo_key == "sa1" &
     spec$property_key == "house" & spec$fe_key == "none") ||
  (spec$outcome_key == "shops1" & spec$geo_key == "sa2" &
     spec$property_key == "apartment" & spec$fe_key == "none")
}

full_runs <- lapply(seq_len(nrow(full_specs)), function(i) {
  spec <- full_specs[i]
  fit_specification(
    spec,
    bootstrap = TRUE,
    keep_draws = representative_distribution(spec),
    seed = BOOTSTRAP_SEED + spec$spec_id
  )
})

paired_results <- rbindlist(lapply(full_runs, `[[`, "summary"), fill = TRUE)
bootstrap_draws <- rbindlist(lapply(full_runs, `[[`, "draws"), fill = TRUE)

if ("p_value_delta" %in% names(paired_results)) {
  paired_results[, global_fdr_q := p.adjust(p_value_delta, method = "BH")]
  paired_results[, within_outcome_fdr_q :=
                   p.adjust(p_value_delta, method = "BH"), by = outcome_key]
}

# Relative changes make the two primary outcomes easier to compare visually.
# Inference remains based on the raw coefficient gap Delta.
paired_results[, relative_ci_a := 100 * ci_low / benchmark]
paired_results[, relative_ci_b := 100 * ci_high / benchmark]
paired_results[, relative_ci_low := pmin(relative_ci_a, relative_ci_b)]
paired_results[, relative_ci_high := pmax(relative_ci_a, relative_ci_b)]
paired_results[, absorption_percent := 100 * absorption_r2]
paired_results[, c("relative_ci_a", "relative_ci_b") := NULL]

setorder(
  paired_results, outcome_order, geo_order, property_order, fe_order
)
paired_results[
  1:min(.N, 12L),
  .(outcome, geography, property_type, fixed_effects, delta,
    bootstrap_se, ci_low, ci_high, p_value_delta)
]
```

# Measurement-error diagnostics

The diagnostic table describes the spatial regressor before conditioning on
the hedonic controls. Its scale-free error measure is

\[
\operatorname{RMSE/SD(exact)}
=
\frac{
\sqrt{N^{-1}\sum_i(\widetilde S_i^g-S_i)^2}
}{
\operatorname{SD}(S_i)
}.
\]

A value of one means that the proxy error is as large as one standard
deviation of the exact transformed variable. The variance ratio and
correlations describe the error structure; they are not separate tests of
classical or Berkson error.

```{r measurement-error-diagnostics}
diagnostic_specs <- cross_join(cross_join(outcomes, geographies), property_types)
setorder(diagnostic_specs, outcome_order, geo_order, property_order)

measurement_diagnostics <- rbindlist(lapply(seq_len(nrow(diagnostic_specs)), function(i) {
  spec <- diagnostic_specs[i]
  proxy_var <- paste0(spec$stem, "_", spec$suffix)
  data <- transactions[house == spec$house_value, c(spec$stem, proxy_var), with = FALSE]
  data <- data[complete.cases(data)]
  exact_raw <- data[[spec$stem]]
  proxy_raw <- data[[proxy_var]]
  exact <- transform_spatial(exact_raw, spec$stem)
  proxy <- transform_spatial(proxy_raw, spec$stem)
  error <- proxy - exact

  cbind(
    spec_metadata(cbind(spec, fixed_effects[1L]))[, !c("fe_key", "fixed_effects", "fe_order", "sample_key")],
    data.table(
      mean_proxy_error_raw = mean(proxy_raw - exact_raw),
      rmse_sd_exact = sqrt(mean(error^2)) / sd(exact),
      exact_proxy_variance_ratio = var(exact) / var(proxy),
      corr_proxy_error = cor(proxy, error),
      corr_exact_error = cor(exact, error),
      N = length(exact)
    )
  )
}), fill = TRUE)
```

# Exact FWL decomposition

The exact and centroid models use the same complete-case observations,
controls, and fixed effects. After applying this common residualisation
operator, write the residuals as \(y\), \(s\), and
\(\widetilde s=s+e\). The Frisch--Waugh--Lovell theorem gives

\[
\widehat\beta_C
=
\frac{\operatorname{Cov}(s+e,y)}
{\operatorname{Var}(\widetilde s)},
\qquad
\widehat\beta_E
=
\frac{\operatorname{Cov}(s,y)}
{\operatorname{Var}(s)}.
\]

Expanding the first covariance and subtracting the second coefficient gives
the exact fitted coefficient-gap accounting

\[
\widehat\Delta
=
\frac{\operatorname{Cov}(e,y)}
{\operatorname{Var}(\widetilde s)}
+
\operatorname{Cov}(s,y)
\left[
\frac{1}{\operatorname{Var}(\widetilde s)}
-
\frac{1}{\operatorname{Var}(s)}
\right].
\]

The first term is the partial error--outcome covariance component. The second
is the conditional variance-rescaling component. The identity requires no
classical- or Berkson-error assumption and accounts for the observed
coefficient gap rather than causal bias.

```{r fwl-decomposition}
fwl_one <- function(spec) {
  prepared <- prepare_pair_data(spec)
  data <- prepared$data
  controls <- controls_for(spec$stem)
  # make_formula expects the first RHS component separately; build the three
  # auxiliary formulas explicitly to keep the control set identical.
  rhs <- paste(controls, collapse = " + ")
  formula_for <- function(lhs) {
    if (nzchar(spec$fe_rhs)) {
      as.formula(paste(lhs, "~", rhs, "|", spec$fe_rhs))
    } else {
      as.formula(paste(lhs, "~", rhs))
    }
  }

  model_y <- feols(formula_for("log_price"), data, vcov = "iid",
                   notes = FALSE, warn = FALSE)
  model_exact <- feols(formula_for(".exact"), data, vcov = "iid",
                       notes = FALSE, warn = FALSE)
  model_proxy <- feols(formula_for(".proxy"), data, vcov = "iid",
                       notes = FALSE, warn = FALSE)

  y <- resid(model_y)
  exact <- resid(model_exact)
  proxy <- resid(model_proxy)
  error <- proxy - exact
  var_exact <- var(exact)
  var_proxy <- var(proxy)
  cov_exact_y <- cov(exact, y)
  cov_error_y <- cov(error, y)
  beta_exact <- cov_exact_y / var_exact
  beta_proxy <- cov(proxy, y) / var_proxy
  covariance_component <- cov_error_y / var_proxy
  variance_component <-
    cov_exact_y * (1 / var_proxy - 1 / var_exact)
  delta <- beta_proxy - beta_exact

  data.table(
    benchmark = beta_exact,
    centroid = beta_proxy,
    delta = delta,
    covariance_component = covariance_component,
    variance_component = variance_component,
    decomposition_sum = covariance_component + variance_component,
    identity_error = delta - covariance_component - variance_component,
    conditional_variance_ratio = var_exact / var_proxy,
    conditional_error_variance = var(error),
    conditional_cov_exact_error = cov(exact, error),
    conditional_cov_error_y = cov_error_y,
    N = length(y)
  )
}

fwl_specs <- full_specs[outcome_key %chin% c("cbd", "shops1")]
fwl_results <- rbindlist(lapply(seq_len(nrow(fwl_specs)), function(i) {
  spec <- fwl_specs[i]
  cbind(spec_metadata(spec), fwl_one(spec))
}), fill = TRUE)
```

# Address-weighted centroid evaluation

```{r corrected-centroid-results}
corrected_one <- function(spec) {
  prepared <- prepare_pair_data(spec, corrected = TRUE)
  data <- prepared$data
  model_exact <- feols(
    prepared$formulas$exact, data, vcov = prepared$vcov,
    notes = FALSE, warn = FALSE
  )
  model_proxy <- feols(
    prepared$formulas$proxy, data, vcov = prepared$vcov,
    notes = FALSE, warn = FALSE
  )
  model_weighted <- feols(
    prepared$formulas$weighted, data, vcov = prepared$vcov,
    notes = FALSE, warn = FALSE
  )
  beta_exact <- coefficient_row(model_exact, ".exact")
  beta_proxy <- coefficient_row(model_proxy, ".proxy")
  beta_weighted <- coefficient_row(model_weighted, ".weighted")

  exact_raw <- data[[prepared$stem]]
  proxy_raw <- data[[prepared$proxy_var]]
  weighted_raw <- data[[prepared$corrected_var]]
  rmse_proxy <- sqrt(mean((data$.proxy - data$.exact)^2))
  rmse_weighted <- sqrt(mean((data$.weighted - data$.exact)^2))

  data.table(
    benchmark = beta_exact[["estimate"]],
    benchmark_se = beta_exact[["std.error"]],
    centroid = beta_proxy[["estimate"]],
    centroid_se = beta_proxy[["std.error"]],
    corrected = beta_weighted[["estimate"]],
    corrected_se = beta_weighted[["std.error"]],
    delta_centroid =
      beta_proxy[["estimate"]] - beta_exact[["estimate"]],
    delta_corrected =
      beta_weighted[["estimate"]] - beta_exact[["estimate"]],
    mean_error_centroid = mean(proxy_raw - exact_raw),
    mean_error_corrected = mean(weighted_raw - exact_raw),
    rmse_sd_exact_centroid = rmse_proxy / sd(data$.exact),
    rmse_sd_exact_corrected = rmse_weighted / sd(data$.exact),
    rmse_improvement_percent =
      100 * (rmse_proxy - rmse_weighted) / rmse_proxy,
    N = nobs(model_exact)
  )
}

corrected_specs <- full_specs[outcome_key %chin% c("cbd", "shops1")]
corrected_results <- rbindlist(lapply(seq_len(nrow(corrected_specs)), function(i) {
  spec <- corrected_specs[i]
  cbind(spec_metadata(spec), corrected_one(spec))
}), fill = TRUE)
```

# Greater Sydney and high-risk regional subsamples

```{r subsample-results}
subsample_results <- data.table()
subsample_specs <- cross_join(
  full_specs[outcome_key %chin% c("cbd", "shops1")],
  sample_definitions[sample_key != "full"]
)
setorder(
  subsample_specs, outcome_order, geo_order, sample_order,
  property_order, fe_order
)

subsample_results <- rbindlist(lapply(seq_len(nrow(subsample_specs)), function(i) {
  spec <- subsample_specs[i]
  fit <- fit_specification(spec, sample_key = spec$sample_key)
  fit$summary[, sample := spec$sample]
  fit$summary[, sample_order := spec$sample_order]
  fit$summary
}), fill = TRUE)

full_core_for_samples <- merge(
  paired_results[outcome_key %chin% c("cbd", "shops1")],
  sample_definitions[sample_key == "full", .(sample_key, sample, sample_order)],
  by = "sample_key",
  all.x = TRUE
)
subsample_plot_data <- rbindlist(
  list(full_core_for_samples, subsample_results), fill = TRUE
)
```

# Sydney and non-Sydney descriptive table

```{r regional-descriptives}
transaction_region_summary <- rbindlist(lapply(
  c("Purchase.price", "dist_cbd_km", "shops_1km"),
  function(var) {
    transactions[, .(
      mean = as.numeric(mean(get(var), na.rm = TRUE)),
      median = as.numeric(median(get(var), na.rm = TRUE)),
      sd = as.numeric(sd(get(var), na.rm = TRUE))
    ), by = .(region = fifelse(sa1_sydney, "Sydney", "Non-Sydney"))][
      , variable := var
    ]
  }
))

regional_attribute_summary <- rbindlist(lapply(seq_len(nrow(geographies)), function(i) {
  geo <- geographies[i]
  prefix <- geo$attribute_prefix
  id <- geo$id_var
  vars <- paste0(prefix, "_", c("area_sa", "shr_resid", "HS_stat_A", "var_circ"))
  units <- unique(transactions[, c(id, paste0(prefix, "_sydney"), vars), with = FALSE], by = id)
  rbindlist(lapply(vars, function(var) {
    units[, value := if (endsWith(var, "area_sa")) get(var) / 10000 else get(var)]
    units[!is.na(get(paste0(prefix, "_sydney"))), .(
      mean = as.numeric(mean(value, na.rm = TRUE)),
      median = as.numeric(median(value, na.rm = TRUE)),
      sd = as.numeric(sd(value, na.rm = TRUE))
    ), by = .(region = fifelse(get(paste0(prefix, "_sydney")), "Sydney", "Non-Sydney"))][
      , `:=`(variable = sub(paste0("^", prefix, "_"), "", var), geography = geo$geography)
    ]
  }))
}), fill = TRUE)
```

# Compact publication summaries

```{r publication-summaries}
additional_variable_summary <- paired_results[family == "additional", {
  largest <- which.max(abs(delta))
  .(
    maximum_absolute_gap = abs(delta[largest]),
    largest_gap_specification =
      paste(property_type[largest], fixed_effects[largest], sep = ", "),
    intervals_excluding_zero =
      sum(ci_low > 0 | ci_high < 0),
    adjusted_gaps = sum(within_outcome_fdr_q < 0.05),
    sign_reversals = sum(sign(benchmark) != sign(centroid))
  )
}, by = .(outcome_order, outcome, geo_order, geography)]
setorder(additional_variable_summary, outcome_order, geo_order)

fwl_publication <- copy(fwl_results)
setorder(
  fwl_publication, outcome_order, geo_order, property_order, fe_order
)

weighted_regressor_summary <- corrected_results[
  fe_key == "none",
  .(
    outcome_order, outcome, geo_order, geography,
    property_order, property_type,
    mean_error_centroid, mean_error_corrected,
    rmse_sd_exact_centroid, rmse_sd_exact_corrected,
    rmse_improvement_percent
  )
]
setorder(
  weighted_regressor_summary,
  outcome_order, geo_order, property_order
)
```

# Transaction map

The dated v10 analysis layer retains the transaction coordinates. This chunk
checks for point geometry and creates an offline map over the official ABS
2021 NSW state boundary, without an API dependency. If a coordinate-bearing
source is unavailable, it copies the existing
manuscript map into the generated-output tree and records the fallback
explicitly in `map_status.csv`.

```{r transaction-map}
map_status <- data.table(status = "not run", source = NA_character_, note = NA_character_)

find_point_source <- function(paths) {
  for (path in paths[file.exists(paths)]) {
    layer <- sf::st_layers(path)$name[[1L]]
    layer_sql <- gsub('"', '""', layer, fixed = TRUE)
    probe <- tryCatch(
      sf::st_read(
        path,
        query = sprintf('SELECT * FROM "%s" LIMIT 1', layer_sql),
        quiet = TRUE
      ),
      error = function(e) NULL
    )
    if (!is.null(probe) && inherits(probe, "sf") && nrow(probe) &&
        any(grepl("POINT", as.character(sf::st_geometry_type(probe))))) {
      return(path)
    }
  }
  NA_character_
}

map_candidates <- c(
    file.path(
      PROJECT_ROOT, "Data", "Intermediate",
      "transactions_with_GNAF_and_amenities_2022_2025_revision.gpkg"
    ),
    DATA_FILE
)
point_source <- find_point_source(map_candidates)
state_boundary_path <- file.path(
  PROJECT_ROOT, "Data", "Inputs", "Boundaries",
  "STE_2021_AUST_GDA2020.shp"
)
map_png <- file.path(FIGURE_DIR, "transaction_map.png")
map_pdf <- file.path(FIGURE_DIR, "transaction_map.pdf")

if (!is.na(point_source)) {
    nsw_boundary <- sf::st_read(state_boundary_path, quiet = TRUE)
    nsw_boundary <- nsw_boundary[nsw_boundary$STE_CODE21 == "1", ]
    nsw_boundary <- sf::st_transform(nsw_boundary, 4326)

    points <- sf::st_read(point_source, quiet = TRUE)
    points <- points[
      !is.na(points$Purchase.price) & !is.na(points$Area) &
        points$Purchase.price >= 200000 & points$Purchase.price < 5000000 &
        points$Area < 10000,
    ]
    points <- points[!is.na(points$Property.unit.number), ]
    points$property_type <- factor(
      ifelse(points$Property.unit.number == "", "(i) House", "(ii) Apartment"),
      levels = c("(i) House", "(ii) Apartment")
    )
    points <- sf::st_transform(points, 4326)
    map_plot <- ggplot() +
      geom_sf(
        data = nsw_boundary,
        fill = "#f2f2f2", colour = "#3f3f3f", linewidth = 0.45
      ) +
      geom_sf(
        data = points,
        colour = "#b2182b", size = 0.055, alpha = 0.18
      ) +
      facet_wrap(~property_type, nrow = 1) +
      coord_sf(
        xlim = c(140.8, 154.0), ylim = c(-37.8, -27.8),
        datum = sf::st_crs(4326), expand = FALSE
      ) +
      labs(x = "Longitude", y = "Latitude") +
      theme_minimal(base_size = 10) +
      theme(
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(colour = "#e6e6e6", linewidth = 0.3),
        panel.background = element_rect(fill = "white", colour = NA),
        panel.border = element_rect(colour = "#666666", fill = NA),
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "white", colour = NA)
      )
    ggsave(map_png, map_plot, width = 10, height = 6, dpi = 300)
    ggsave(map_pdf, map_plot, width = 10, height = 6)
    file.copy(map_png, file.path(FIGURE_DIR, "map_300dpi.png"), overwrite = TRUE)
    map_status <- data.table(
      status = "generated from coordinates", source = point_source,
      note = paste(
        "Offline ggplot2/sf map over the ABS 2021 NSW state boundary;",
        "no web-map API required."
      )
    )
} else {
    legacy_map <- file.path(
      PROJECT_ROOT, "Data", "Inputs", "transaction_map_reference.png"
    )
    if (!file.exists(legacy_map)) stop("No coordinate source or legacy map is available")
    file.copy(legacy_map, map_png, overwrite = TRUE)
    file.copy(legacy_map, file.path(FIGURE_DIR, "map_300dpi.png"), overwrite = TRUE)
    map_status <- data.table(
      status = "legacy fallback", source = legacy_map,
      note = paste(
        "The analysis GPKG has no geometry. Re-run Data Preparation while",
        "retaining geometry to regenerate this figure from coordinates."
      )
    )
}
fwrite(map_status, file.path(DATA_DIR, "map_status.csv"))
```

# Publication figures

```{r publication-figures}
publication_theme <- theme_minimal(base_size = 10) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    plot.title.position = "plot"
  )

save_publication_plot <- function(plot, stem, width = 9, height = 6) {
  ggsave(file.path(FIGURE_DIR, paste0(stem, ".pdf")), plot,
         width = width, height = height)
  ggsave(file.path(FIGURE_DIR, paste0(stem, ".png")), plot,
         width = width, height = height, dpi = 300)
}

core_gap_data <- copy(
  paired_results[outcome_key %chin% c("cbd", "shops1")]
)
core_gap_data[, geography := factor(
  geography, levels = c("SA1", "SA2", "Postcode")
)]
core_gap_data[, property_type := factor(
  property_type, levels = c("House", "Apartment")
)]
core_gap_data[, specification := factor(
  paste(geography, fixed_effects, sep = " - "),
  levels = rev(as.vector(outer(
    c("SA1", "SA2", "Postcode"), c("No FE", "SA3", "SA4"),
    paste, sep = " - "
  )))
)]

core_gap_plot <- ggplot(
  core_gap_data,
  aes(
    x = relative_change_percent,
    y = specification,
    xmin = relative_ci_low,
    xmax = relative_ci_high,
    colour = geography
  )
) +
  geom_vline(xintercept = 0, colour = "grey45", linewidth = 0.45) +
  geom_errorbar(orientation = "y", width = 0.15) +
  geom_point(size = 2) +
  facet_grid(property_type ~ outcome) +
  labs(
    x = "Relative coefficient change (%)",
    y = NULL,
    colour = "Centroid geography",
    caption = paste(
      "Zero means no coefficient change. Negative values generally indicate",
      "attenuation when the benchmark and centroid coefficients have the same sign."
    )
  ) +
  publication_theme
save_publication_plot(core_gap_plot, "core_coefficient_gaps", 10, 7)

radius_data <- copy(
  paired_results[outcome_key %chin% c("shops1", "shops2", "shops5")]
)
radius_data[, geography := factor(
  geography, levels = c("SA1", "SA2", "Postcode")
)]
radius_data[, property_type := factor(
  property_type, levels = c("House", "Apartment")
)]
radius_data[, radius_km := as.numeric(sub("shops", "", outcome_key))]

radius_plot <- ggplot(
  radius_data,
  aes(
    x = radius_km,
    y = delta,
    ymin = ci_low,
    ymax = ci_high,
    colour = fixed_effects,
    group = fixed_effects
  )
) +
  geom_hline(yintercept = 0, colour = "grey45", linewidth = 0.45) +
  geom_errorbar(width = 0.08, alpha = 0.65) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = c(1, 2, 5)) +
  facet_grid(property_type ~ geography, scales = "free_y") +
  labs(
    x = "Shop-count radius (km)",
    y = "Coefficient gap (centroid minus exact)",
    colour = "Fixed effects",
    caption = "Intervals are paired spatial-cluster bootstrap intervals for the coefficient gap."
  ) +
  publication_theme
save_publication_plot(
  radius_plot, "shop_radius_coefficient_gaps", 10, 6.5
)

distribution_keys <- unique(
  bootstrap_draws[, .(outcome_key, geo_key, property_key, fe_key)]
)
distribution_summary <- merge(
  distribution_keys,
  paired_results[, .(
    outcome_key, geo_key, property_key, fe_key,
    outcome, geography, property_type, fixed_effects,
    observed_delta = delta, ci_low, ci_high
  )],
  by = c("outcome_key", "geo_key", "property_key", "fe_key")
)
distribution_summary[, panel := paste(
  outcome, geography, property_type, fixed_effects, sep = " - "
)]
distribution_data <- copy(bootstrap_draws)
distribution_data[, panel := paste(
  outcome, geography, property_type, fixed_effects, sep = " - "
)]

distribution_plot <- ggplot(distribution_data, aes(x = delta)) +
  geom_vline(xintercept = 0, colour = "grey35", linewidth = 0.45) +
  geom_density(fill = "#56B4E9", colour = "#0072B2", alpha = 0.35) +
  geom_vline(
    data = distribution_summary,
    aes(xintercept = observed_delta),
    colour = "#D55E00",
    linewidth = 0.7
  ) +
  geom_vline(
    data = distribution_summary,
    aes(xintercept = ci_low),
    colour = "#D55E00",
    linetype = "dashed",
    linewidth = 0.45
  ) +
  geom_vline(
    data = distribution_summary,
    aes(xintercept = ci_high),
    colour = "#D55E00",
    linetype = "dashed",
    linewidth = 0.45
  ) +
  facet_wrap(~panel, scales = "free", ncol = 2) +
  labs(
    x = "Paired bootstrap coefficient gap",
    y = "Density",
    caption = paste(
      "Black: no difference. Orange: observed gap.",
      "Dashed orange: paired 95% interval."
    )
  ) +
  publication_theme
save_publication_plot(
  distribution_plot, "paired_bootstrap_gap_distributions", 10, 7
)

if (nrow(subsample_plot_data)) {
  make_subsample_plot <- function(data, outcome_value, measure, y_label, stem) {
    plot_data <- data[outcome_key == outcome_value & !is.na(get(measure))]
    plot_data[, sample := factor(sample, levels = sample_definitions$sample)]
    plot_data[, geography := factor(
      geography, levels = c("SA1", "SA2", "Postcode")
    )]
    plot_data[, series := factor(
      paste(property_type, fixed_effects, sep = " - "),
      levels = as.vector(t(outer(
        c("House", "Apartment"), c("No FE", "SA3", "SA4"),
        paste, sep = " - "
      )))
    )]
    plot <- ggplot(
      plot_data,
      aes(x = sample, y = get(measure), fill = series)
    ) +
      geom_hline(yintercept = 0, colour = "grey55", linewidth = 0.35) +
      geom_col(position = position_dodge(width = 0.8), width = 0.72) +
      facet_wrap(~geography, ncol = 1, scales = "free_y") +
      labs(
        x = NULL, y = y_label, fill = NULL,
        caption = "Specifications with no focal-regressor variation are not estimable and are omitted."
      ) +
      publication_theme +
      theme(axis.text.x = element_text(angle = 25, hjust = 1))
    save_publication_plot(plot, stem, 8, 10)
  }

  make_subsample_plot(
    subsample_plot_data, "cbd", "delta",
    "Centroid coefficient minus exact-coordinate coefficient",
    "subsample_cbd_coefficient_gaps"
  )
  make_subsample_plot(
    subsample_plot_data, "cbd", "se_difference",
    "Centroid SE minus benchmark SE",
    "subsample_cbd_standard_error_gaps"
  )
  make_subsample_plot(
    subsample_plot_data, "shops1", "delta",
    "Centroid coefficient minus exact-coordinate coefficient",
    "subsample_shops_coefficient_gaps"
  )
  make_subsample_plot(
    subsample_plot_data, "shops1", "se_difference",
    "Centroid SE minus benchmark SE",
    "subsample_shops_standard_error_gaps"
  )

}
```

# Machine-readable exports

```{r data-exports}
fwrite(sample_cleaning_flow, file.path(DATA_DIR, "sample_cleaning_flow.csv"))
fwrite(ambiguous_match_summary, file.path(DATA_DIR, "ambiguous_match_details.csv"))
fwrite(ambiguous_counts, file.path(DATA_DIR, "ambiguous_matches_summary.csv"))
fwrite(ambiguity_by_property, file.path(DATA_DIR, "ambiguous_matches_by_property.csv"))
fwrite(descriptive_long, file.path(DATA_DIR, "descriptive_statistics_long.csv"))
fwrite(descriptive_table, file.path(DATA_DIR, "descriptive_statistics_wide.csv"))
fwrite(measurement_diagnostics, file.path(DATA_DIR, "measurement_error_diagnostics.csv"))
fwrite(paired_results, file.path(DATA_DIR, "paired_coefficient_results.csv"))
fwrite(
  paired_results[family == "primary"],
  file.path(DATA_DIR, "paired_primary_72_results.csv")
)
fwrite(
  paired_results[family == "additional"],
  file.path(DATA_DIR, "paired_additional_spatial_results.csv")
)
fwrite(bootstrap_draws, file.path(DATA_DIR, "representative_bootstrap_draws.csv"))
fwrite(fwl_results, file.path(DATA_DIR, "fwl_decomposition_results.csv"))
fwrite(corrected_results, file.path(DATA_DIR, "corrected_centroid_results.csv"))
fwrite(
  weighted_regressor_summary,
  file.path(DATA_DIR, "weighted_centroid_regressor_summary.csv")
)
fwrite(
  additional_variable_summary,
  file.path(DATA_DIR, "additional_spatial_variable_summary.csv")
)
if (ncol(subsample_results)) {
  fwrite(subsample_results, file.path(DATA_DIR, "subsample_results.csv"))
} else {
  writeLines("", file.path(DATA_DIR, "subsample_results.csv"))
}
fwrite(transaction_region_summary, file.path(DATA_DIR, "transaction_region_summary.csv"))
fwrite(regional_attribute_summary, file.path(DATA_DIR, "regional_attribute_summary.csv"))
```

# LaTeX tables generated with modelsummary

```{r latex-table-functions}
format_number <- function(x, digits = 4L) {
  ifelse(is.na(x), "", formatC(x, format = "f", digits = digits))
}

format_integer <- function(x) {
  ifelse(is.na(x), "", format(round(x), big.mark = ",", scientific = FALSE))
}

format_model_cell <- function(beta, se) {
  ifelse(
    is.na(beta) | is.na(se), "",
    paste0(format_number(beta), " (", format_number(se), ")")
  )
}

format_interval <- function(low, high) {
  ifelse(
    is.na(low) | is.na(high), "",
    paste0("[", format_number(low), ", ", format_number(high), "]")
  )
}

export_latex_table <- function(data, filename, caption, label, notes = "") {
  tabular <- suppressWarnings(modelsummary::datasummary_df(
    as.data.frame(data), output = "latex_tabular", fmt = 4,
    escape = FALSE
  ))
  output <- c(
    "\\begin{table}[!htbp]",
    "\\centering",
    paste0("\\caption{", caption, "}"),
    paste0("\\label{", label, "}"),
    "\\resizebox{\\textwidth}{!}{%",
    as.character(tabular),
    "}",
    if (nzchar(notes)) paste0(
      "\\par\\smallskip{\\footnotesize\\raggedright\\textit{Notes}: ",
      notes, "\\par}"
    ) else character(),
    "\\end{table}"
  )
  writeLines(output, file.path(TABLE_DIR, filename))
}

main_results_table_data <- function(data) {
  data[, .(
    Geography = geography,
    `Property type` = property_type,
    `Fixed effects` = fixed_effects,
    `Exact coordinate` = format_model_cell(benchmark, benchmark_se),
    `Centroid proxy` = format_model_cell(centroid, centroid_se),
    `Gap $\\Delta$` = format_number(delta),
    `Relative change (\\%)` =
      format_number(relative_change_percent, 1L),
    `Paired 95\\% CI` = format_interval(ci_low, ci_high),
    `FE absorption (\\%)` =
      format_number(absorption_percent, 1L),
    N = format_integer(N)
  )]
}

diagnostic_table_data <- function(data) {
  data[, .(
    `Spatial variable` = outcome,
    Centroid = geography,
    `Property type` = property_type,
    `Mean proxy error` = format_number(mean_proxy_error_raw),
    `RMSE / SD(exact)` = format_number(rmse_sd_exact),
    `Exact/proxy variance` =
      format_number(exact_proxy_variance_ratio),
    `Corr(exact, error)` = format_number(corr_exact_error),
    `Corr(proxy, error)` = format_number(corr_proxy_error),
    N = format_integer(N)
  )]
}

weighted_coefficient_table_data <- function(data) {
  data[, .(
    Geography = geography,
    `Property type` = property_type,
    `Fixed effects` = fixed_effects,
    `Exact coordinate` = format_model_cell(benchmark, benchmark_se),
    `Geometric centroid` = format_model_cell(centroid, centroid_se),
    `Address-weighted centroid` = format_model_cell(corrected, corrected_se),
    `Geometric gap` = format_number(delta_centroid),
    `Weighted gap` = format_number(delta_corrected),
    N = format_integer(N)
  )]
}

weighted_regressor_table_data <- function(data) {
  data[, .(
    `Spatial variable` = outcome,
    Geography = geography,
    `Property type` = property_type,
    `Geometric mean error` = format_number(mean_error_centroid),
    `Weighted mean error` = format_number(mean_error_corrected),
    `Geometric RMSE / SD(exact)` =
      format_number(rmse_sd_exact_centroid),
    `Weighted RMSE / SD(exact)` =
      format_number(rmse_sd_exact_corrected),
    `RMSE improvement (\\%)` =
      format_number(rmse_improvement_percent, 1L)
  )]
}

fwl_table_data <- function(data) {
  data[, .(
    Geography = geography,
    `Property type` = property_type,
    `Fixed effects` = fixed_effects,
    `Coefficient gap` = format_number(delta),
    `Partial error--outcome covariance` =
      format_number(covariance_component),
    `Variance rescaling` = format_number(variance_component),
    `Conditional exact/proxy variance` =
      format_number(conditional_variance_ratio),
    N = format_integer(N)
  )]
}
```

```{r publication-tables}
cleaning_display <- sample_cleaning_flow[
  stage %chin% c(
    "Latest eligible dealing-address sales",
    "Excluded sales with conflicting model inputs",
    "Final unambiguous validation sample",
    "Final houses",
    "Final apartments"
  ),
  .(
    Stage = stage,
    Unit = unit,
    Count = format_integer(count)
  )
]

export_latex_table(
  cleaning_display,
  "sample_cleaning_flow.tex",
  "Construction of the Unique Address-Coordinate Validation Sample",
  "tab:sample-cleaning",
  paste(
    "A sale is identified by dealing number and normalized address.",
    "The latest downloaded record is retained. Rows with identical model",
    "inputs are collapsed, while sales retaining conflicting exact-coordinate",
    "or other model inputs are excluded rather than assigned an arbitrary",
    "G-NAF match."
  )
)

export_latex_table(
  descriptive_table[, .(
    Variable = label,
    `House mean` = ifelse(label %chin% c("Price (AUD)", "Area (sqm)"),
                          format_integer(mean_House), format_number(mean_House, 2L)),
    `House median` = ifelse(label %chin% c("Price (AUD)", "Area (sqm)"),
                            format_integer(median_House), format_number(median_House, 2L)),
    `House SD` = ifelse(label %chin% c("Price (AUD)", "Area (sqm)"),
                        format_integer(sd_House), format_number(sd_House, 2L)),
    `Apartment mean` = ifelse(label %chin% c("Price (AUD)", "Area (sqm)"),
                              format_integer(mean_Apartment), format_number(mean_Apartment, 2L)),
    `Apartment median` = ifelse(label %chin% c("Price (AUD)", "Area (sqm)"),
                                format_integer(median_Apartment), format_number(median_Apartment, 2L)),
    `Apartment SD` = ifelse(label %chin% c("Price (AUD)", "Area (sqm)"),
                            format_integer(sd_Apartment), format_number(sd_Apartment, 2L)),
    Source = source
  )],
  "descriptive_statistics.tex",
  "Descriptive Statistics of NSW Property Transactions (2022--2025)",
  "tab:data-sum",
  paste(
    "Statistics are calculated for unique dealing-address sales after the",
    "price and land-area restrictions and exclusion of conflicting",
    "address-coordinate matches."
  )
)

export_latex_table(
  diagnostic_table_data(
    measurement_diagnostics[outcome_key %chin% c("cbd", "shops1")]
  ),
  "measurement_error_diagnostics.tex",
  "Measurement-Error Diagnostics for the Primary Spatial Variables",
  "tab:me-diagnostics",
  paste(
    "Error is centroid proxy minus exact-coordinate value.",
    "Mean error uses raw units. The other columns use the regression",
    "transformation. RMSE / SD(exact) equals one when proxy RMSE is as",
    "large as one standard deviation of the exact variable.",
    "The diagnostics describe error structure; they are not formal tests",
    "of classical or Berkson measurement error."
  )
)

table_labels <- list(
  cbd = c("CBD Distance", "cbd"),
  shops1 = c("Shops within 1 km", "shop")
)

for (outcome_key_value in names(table_labels)) {
  outcome_title <- table_labels[[outcome_key_value]][[1L]]
  label_stem <- table_labels[[outcome_key_value]][[2L]]
  outcome_data <- paired_results[outcome_key == outcome_key_value]

  export_latex_table(
    main_results_table_data(outcome_data),
    paste0("core_", outcome_key_value, "_results.tex"),
    paste0(outcome_title, ": Exact and Centroid-Based Results"),
    paste0("tab:core-", label_stem),
    paste(
      "$\\Delta=\\widehat\\beta_{centroid}-",
      "\\widehat\\beta_{exact}$.",
      "Relative change is",
      "$100(\\widehat\\beta_{centroid}/",
      "\\widehat\\beta_{exact}-1)$.",
      "Coefficient-specific standard errors are in parentheses.",
      "The confidence interval is for $\\Delta$ and uses 999 paired",
      "one-way spatial-cluster multiplier draws.",
      "FE absorption is the percentage of proxy variation jointly",
      "absorbed by the stated spatial and month-by-year fixed effects."
    )
  )
}

export_latex_table(
  weighted_regressor_table_data(weighted_regressor_summary),
  "weighted_centroid_regressor_quality.tex",
  "Regressor Quality of Geometric and Address-Weighted Centroids",
  "tab:weighted-regressor-quality",
  paste(
    "Mean errors use raw units; RMSE measures use the regression",
    "transformation. Positive improvement means that the address-weighted",
    "proxy has lower RMSE than the geometric centroid proxy."
  )
)

for (outcome_key_value in c("cbd", "shops1")) {
  outcome_title <- outcomes[outcome_key == outcome_key_value, outcome]
  label_stem <- if (outcome_key_value == "cbd") "cbd" else "shop"
  outcome_data <- corrected_results[outcome_key == outcome_key_value]

  export_latex_table(
    weighted_coefficient_table_data(outcome_data),
    paste0("weighted_centroid_", outcome_key_value, "_results.tex"),
    paste0(outcome_title, ": Address-Weighted Centroid Coefficients"),
    paste0("tab:weighted-", label_stem),
    paste(
      "Both gaps are relative to the exact-coordinate coefficient.",
      "All three regressors are evaluated on the same sample."
    )
  )
}

additional_display <- additional_variable_summary[, .(
  `Spatial variable` = outcome,
  Geography = geography,
  `Largest absolute gap` = format_number(maximum_absolute_gap),
  `Largest-gap specification` = largest_gap_specification,
  `95\\% intervals excluding zero` =
    paste0(intervals_excluding_zero, " / 6"),
  `BH-adjusted gaps` = paste0(adjusted_gaps, " / 6"),
  `Sign reversals` = paste0(sign_reversals, " / 6")
)]

export_latex_table(
  additional_display,
  "additional_spatial_variables_summary.tex",
  "Centroid Sensitivity of Additional Spatial Variables",
  "tab:additional-spatial-summary",
  paste(
    "Each row summarizes two property types and three fixed-effect",
    "specifications. Gaps are centroid minus exact-coordinate coefficients.",
    "BH adjustment is applied within each spatial variable across all",
    "18 geography--property--fixed-effect comparisons."
  )
)

fwl_table_labels <- list(
  cbd = c("CBD Distance", "cbd"),
  shops1 = c("Shops within 1 km", "shops")
)

for (outcome_key_value in names(fwl_table_labels)) {
  outcome_title <- fwl_table_labels[[outcome_key_value]][[1L]]
  label_stem <- fwl_table_labels[[outcome_key_value]][[2L]]
  outcome_data <- fwl_publication[outcome_key == outcome_key_value]

  export_latex_table(
    fwl_table_data(outcome_data),
    paste0("fwl_", outcome_key_value, "_decomposition.tex"),
    paste0("Complete Coefficient-Gap Accounting: ", outcome_title),
    paste0("tab:fwl-", label_stem),
    paste(
      "The underlying unrounded partial error--outcome covariance and",
      "variance-rescaling components sum exactly to the reported",
      "coefficient gap after partialling out the controls and fixed effects.",
      "The identity is descriptive and is not a causal decomposition."
    )
  )
}

regional_display <- rbindlist(list(
  transaction_region_summary[, .(
    Variable = variable, Geography = "Transaction",
    Region = region, Mean = format_number(mean, 2L),
    Median = format_number(median, 2L), SD = format_number(sd, 2L)
  )],
  regional_attribute_summary[, .(
    Variable = variable, Geography = geography,
    Region = region, Mean = format_number(mean, 2L),
    Median = format_number(median, 2L), SD = format_number(sd, 2L)
  )]
), fill = TRUE)
regional_labels <- c(
  Purchase.price = "Price (AUD)",
  dist_cbd_km = "CBD distance (km)",
  shops_1km = "Shops within 1 km",
  area_sa = "Area (hectares)",
  shr_resid = "Residential share",
  HS_stat_A = "Clustering statistic",
  var_circ = "Boundary irregularity"
)
regional_display[, Variable := unname(regional_labels[Variable])]

export_latex_table(
  regional_display,
  "regional_descriptive_statistics.tex",
  "Sydney and Non-Sydney Transactions and Geographic Units",
  "tab:data_summary_syd",
  "Geographic attributes are summarized over unique units rather than transaction-weighted repeated rows."
)
```

# Validation, bundle, and manifest

```{r validation-and-bundle}
expected_all_pairs <- nrow(outcomes) * 3L * 2L * 3L
expected_subsamples <- 2L * 3L * 5L * 2L * 3L

validation <- data.table(
  check = c(
    "clean sample size", "house sample size", "apartment sample size",
    "all paired rows", "FWL rows", "FWL identity",
    "corrected rows", "diagnostic rows", "subsample rows",
    "subsample non-estimable rows"
  ),
  expected = c(
    340856,
    281621,
    59235,
    expected_all_pairs,
    36,
    1e-8,
    36,
    nrow(outcomes) * 3L * 2L,
    expected_subsamples,
    NA
  ),
  observed = c(
    nrow(transactions),
    transactions[house == TRUE, .N],
    transactions[house == FALSE, .N],
    nrow(paired_results),
    nrow(fwl_results),
    max(abs(fwl_results$identity_error), na.rm = TRUE),
    nrow(corrected_results),
    nrow(measurement_diagnostics),
    nrow(subsample_results),
    sum(grepl("^not estimable", subsample_results$status))
  )
)
validation[, pass := ifelse(
  is.na(expected), TRUE,
  ifelse(check == "FWL identity", observed < expected, observed == expected)
)]
fwrite(validation, file.path(DATA_DIR, "validation_checks.csv"))

run_config <- data.table(
  run_time = as.character(Sys.time()),
  bootstrap_reps = BOOTSTRAP_REPS,
  bootstrap_seed = BOOTSTRAP_SEED,
  transaction_rows = nrow(transactions),
  data_file = DATA_FILE
)
fwrite(run_config, file.path(DATA_DIR, "run_configuration.csv"))

results_bundle <- list(
  configuration = run_config,
  validation = validation,
  sample_cleaning_flow = sample_cleaning_flow,
  ambiguous_match_summary = ambiguous_match_summary,
  ambiguity_by_property = ambiguity_by_property,
  descriptive_statistics = descriptive_long,
  measurement_diagnostics = measurement_diagnostics,
  paired_results = paired_results,
  representative_bootstrap_draws = bootstrap_draws,
  fwl_results = fwl_results,
  fwl_publication = fwl_publication,
  corrected_results = corrected_results,
  weighted_regressor_summary = weighted_regressor_summary,
  additional_variable_summary = additional_variable_summary,
  subsample_results = subsample_results,
  transaction_region_summary = transaction_region_summary,
  regional_attribute_summary = regional_attribute_summary,
  map_status = map_status
)
saveRDS(results_bundle, file.path(OUTPUT_DIR, "v10_results.rds"), compress = "xz")
package_versions <- vapply(
  packages_used,
  function(package) as.character(utils::packageVersion(package)),
  FUN.VALUE = character(1)
)
writeLines(
  c(
    R.version.string,
    paste(names(package_versions), package_versions, sep = ": ")
  ),
  file.path(OUTPUT_DIR, "session-info.txt")
)

all_output_files <- list.files(OUTPUT_DIR, recursive = TRUE, full.names = TRUE)
output_prefix_length <- nchar(normalizePath(OUTPUT_DIR, mustWork = TRUE)) + 2L
manifest <- data.table(
  file = substring(normalizePath(all_output_files, mustWork = TRUE), output_prefix_length),
  bytes = file.info(all_output_files)$size,
  md5 = unname(tools::md5sum(all_output_files))
)
manifest <- manifest[file != "manifest.csv"]
fwrite(manifest, file.path(OUTPUT_DIR, "manifest.csv"))

validation
```

# Run interpretation

The manuscript and response letter must be updated only from the exported
objects. In particular:

- coefficient-specific standard errors do not provide inference for
  \(\Delta\);
- the paired bootstrap is a one-way spatial-cluster influence-function
  multiplier bootstrap, assumes independence across released-geography
  clusters after conditioning, and does not re-estimate every model in every
  draw;
- RMSE / SD(exact), variance ratios, and correlations are diagnostics rather than tests
  that separately identify classical and Berkson errors;
- a positive address-weighted RMSE improvement is direct evidence that the
  correction improves the regressor, even if a particular price coefficient
  happens to move toward the benchmark.
